Loss of CTRP10 results in female obesity with preserved metabolic health

Version 1

Abstract

Abstract

Obesity is a major risk factor for type 2 diabetes, dyslipidemia, cardiovascular disease, and hypertension. Intriguingly, there is a subset of metabolically healthy obese (MHO) individuals who are seemingly able to maintain a healthy metabolic profile free of metabolic syndrome. The molecular underpinnings of MHO, however, are not well understood. Here, we report that CTRP10/C1QL2-deficient mice represent a unique female model of MHO. CTRP10 modulates weight gain in a striking and sexually dimorphic manner. Female, but not male, mice lacking CTRP10 develop obesity with age on a low-fat diet while maintaining an otherwise healthy metabolic profile. When fed an obesogenic diet, female Ctrp10 knockout (KO) mice show rapid weight gain. Despite pronounced obesity, Ctrp10 KO female mice do not develop steatosis, dyslipidemia, glucose intolerance, insulin resistance, oxidative stress, or low-grade inflammation. Obesity is largely uncoupled from metabolic dysregulation in female KO mice. Multi-tissue transcriptomic analyses highlighted gene expression changes and pathways associated with insulin-sensitive obesity. Transcriptional correlation of the differentially expressed gene (DEG) orthologous in humans also show sex differences in gene connectivity within and across metabolic tissues, underscoring the conserved sex-dependent function of CTRP10. Collectively, our findings suggest that CTRP10 negatively regulates body weight in females, and that loss of CTRP10 results in benign obesity with largely preserved insulin sensitivity and metabolic health. This female MHO mouse model is valuable for understanding sex-biased mechanisms that uncouple obesity from metabolic dysfunction.

Introduction

The prevalence of obesity has nearly tripled in the past four decades and the underlying cause is complex and multifactorial (1, 2). Genetics, environmental and social factors, and demographics all play a contributing role in contributing to excessive weight gain in the setting of overnutrition (3, 4). Although obesity is a major risk factor for type 2 diabetes, dyslipidemia, cardiovascular disease, and hypertension, not all obese individuals develop the metabolic syndrome (5). There is a subset of metabolically healthy obese (MHO) individuals with an apparently healthy metabolic profile free of some or most components of the metabolic syndrome (6, 7). The molecular and physiological underpinnings of MHO are, however, not well understood. Novel preclinical animal models that can recapitulate features of MHO will be valuable in illuminating pathways that resist the deleterious effects of obesity and provide new therapeutic avenues to mitigate obesity-linked comorbidities.

The mechanisms that normally maintain body weight and metabolic homeostasis are complex and involve both cell autonomous and non-cell autonomous mechanisms. Tissue crosstalk mediated by paracrine and endocrine factors plays an especially important role in coordinating metabolic processes across organ systems to maintain energy balance (8). Of the secretory proteins that circulate in plasma, C1q/TNF-related proteins (CTRP1-15) have emerged as important regulators of insulin sensitivity, and glucose and lipid metabolism (9). We originally identified the first seven members of the CTRP family based on shared sequence homology to the insulin-sensitizing adipokine, adiponectin (10), and subsequently characterized additional members (11-16). All fifteen CTRPs share a common C-terminal globular C1q domain and are part of the much larger C1q family (17, 18). The use of gain- and loss-of- function mouse models has helped establish CTRP’s role in controlling various aspects of sugar and fat metabolism (12, 19-34). Additional diverse functions of CTRPs have also been demonstrated in the cardiovascular (35-47), renal (48, 49), immune (20, 50, 51), sensory (52, 53), gastrointestinal (54), musculoskeletal (55-57), and the nervous system (58-61).

Of the family members, CTRP10 (also known as C1QL2) is understudied and consequently only limited information is available concerning its function. The best characterized role of CTRP10 is in the central nervous system (CNS). It has been shown that CTRP10 secreted from mossy fibers is required for the proper clustering of kainite-type glutamate receptors on postsynaptic CA3 pyramidal neurons in the hippocampus (62). It serves as a transsynaptic organizer by directly binding to neurexin3 (Nrx3) on the presynaptic terminals, and to GluK2 and GluK4 on the postsynaptic terminals (62). Additional putative roles of CTRP10 in the CNS have also been suggested. Genome-wide association studies (GWAS) have implicated CTRP10/C1QL2 in cocaine use disorder (63). In rat models of depression, Ctrp10 expression is increased in the dentate gyrus and reduced in the nucleus accumbens (64). In humans with a history of psychiatric disorders (e.g., schizophrenia), the expression of CTRP10 is elevated in the dorsolateral prefrontal cortex of both males and females (64). Whether and how CTRP10 contributes to addictive behavior and psychiatric disorders is unknown.

The potential function of CTRP10 in peripheral tissues, however, is essentially unknown and unexplored. The present study was motivated by the well documented metabolic functions of many CTRP family members we have characterized to date using genetic loss-of-function mouse models (19-28, 30, 31). We determined that the expression of Ctrp10 in peripheral tissues is modulated by diet and nutritional states, and thus may have a metabolic role. We therefore used a genetic loss-of-function mouse model to determine if CTRP10 is required for regulating systemic metabolism. We unexpectedly discovered a female-specific requirement of CTRP10 for body weight control. We showed that the Ctrp10 KO mice represent a unique female model of MHO with largely preserved insulin sensitivity and metabolic health. This valuable mouse model can be used to inform sex-dependent mechanisms that uncouple obesity from insulin resistance, dyslipidemia, and metabolic dysfunction.

Results

Nutritional regulation of Ctrp10 expression in the brain and peripheral tissues

CTRP10 protein is highly conserved from zebrafish to human (Fig. 1A), with amino acid identity of 67%, 71%, 77%, and 94% between the full-length human protein and the fish, frog, chicken, and mouse orthologs, respectively. The conservation is much higher at the C-terminal globular C1q domain (93-100% identity) between the orthologs. Among the 12 different mouse tissue examined, brain had the highest expression of Ctrp10 (Fig. 1B), consistent with previous findings (65). Expression of Ctrp10 in peripheral tissues was variable and generally much lower than in the brain (Fig. 1B). We first determined whether Ctrp10 expression is modulated by nutrition and metabolic state. Male mice were subjected to fasting and refeeding. In the refed period after an overnight fast, we observed a significant downregulation of Ctrp10 in the visceral (gonadal) white adipose tissue (gWAT), liver, skeletal muscle, kidney, cerebellum, cortex, and hypothalamus relative to the fasted state (Fig. 1C). Next, we examined whether an obesogenic diet alters the expression of Ctrp10. Male mice fed a high-fat diet for 12 weeks had a modest increase in Ctrp10 expression in brown adipose tissue (BAT) and heart and decreased expression in skeletal muscle relative to mice fed a control low-fat diet (LFD) (Fig. 1D). These data indicate that Ctrp10 expression is dynamically regulated by acute alterations in energy balance, and perhaps to a lesser extent in alterations to chronic nutritional state on an obesogenic diet.

Nutritional regulation of Ctrp10 expression.

(A) Sequence alignment of full-length human (GenBank # NP_872334), mouse (NP_997116), chicken (XP_046777733), xenopus frog (XP_031749381), and zebrafish (XP_001920705) CTRP10/C1ql2 using Clustal-omega (133). Identical amino acids are shaded black and similar amino acids are shaded grey. Gaps are indicated by dash lines. Signal peptide, collagen domain with characteristic Gly-X-Y repeats, and the C-terminal globular C1q domain are indicated. (B) Ctrp10 expression across different mouse tissues (n = 10). (C) Expression of Ctrp10 across mouse tissues in response to an overnight (16 h) fast or fasting followed by 2 h refeeding. (D) Expression of Ctrp10 across mouse tissues in response to a high-fat diet (HFD) for 12 weeks or a control low-fat diet (LFD). (E) Generation of Ctrp10 knockout (KO) mice. The entire protein coding region in exon 1 and 2 of Ctrp10 was deleted using CRISPR/Cas9 method and confirmed with DNA sequencing. (F) Wild-type (WT) and KO alleles were confirmed by PCR genotyping. (G) The complete loss of Ctrp10 transcript in KO mice was confirmed in mouse cortex, one of the tissues with high Ctrp10 expression (WT, n = 5; KO, n = 5). All expression levels were normalized to β-actin. All data are presented as mean ± S.E.M. * P < 0.05; ** P < 0.01; *** P < 0.001.

Generation of Ctrp10 knockout (KO) mice

We used mice lacking CTRP10 to address whether this secreted protein has a metabolic role in vivo. The mouse Ctrp10 gene consists of two exons (Fig. 1E). The CRISPR-Cas9 method was used to remove the entire protein coding region spanning exon 1 and 2, thus ensuring a complete null allele (Fig. 1E-F). The targeted allele was confirmed by sequencing. As expected, based on the gene deletion strategy, the Ctrp10 transcript was absent from KO mice (Fig. 1G).

CTRP10 is largely dispensable for metabolic homeostasis in young mice fed a control low-fat diet

The body weight and body composition of male mice fed a LFD were not different between genotypes (Fig. 2A-B). By 20 weeks of age, female KO mice fed LFD had a modestly higher body weight relative to WT controls (Fig. 2 C), though the body composition was not different between genotypes (Fig. 2D). Food intake, physical activity, and energy expenditure as measured by indirect calorimetry were also not different between genotypes of either sex across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (Fig. 2E-J). Because Ctrp10 expression is regulated by nutritional states (Fig. 1C), we assessed serum metabolite levels in WT and KO mice in response to fasting and refeeding. No significant differences in fasting and refeeding blood glucose, serum insulin, triglyceride, cholesterol, non-esterified free fatty acids (NEFA), and β-hydroxybutyrate levels were observed between genotypes of either sex, except the female KO mice had slightly lower fasting β-hydroxybutyrate levels (Fig. 3A-B). We performed glucose and insulin tolerance tests to determine any potential differences in glucose handling capacity and insulin sensitivity. No significant differences in glucose and insulin tolerance were noted between genotypes of either sex (Fig. 3C-F). Together, these data indicate that CTRP10 is dispensable for metabolic homeostasis when mice are young (< 20 weeks old) and fed a LFD.

Ctrp10-KO mice fed a low-fat diet have normal body weight and energy balance.

(A-B) Body weight (A) and body composition analysis (B) of fat mass, % fat mass (relative to body weight), lean mass, and % lean mass of WT (n = 17) and KO (n = 14) male mice at 18 weeks of age. (C-D) Body weight (C) and body composition analysis (D) of fat mass, % fat mass (relative to body weight), lean mass, and % lean mass of WT (n = 9) and KO (n = 6) female mice at 13 weeks of age. (E-G) Food intake, physical activity, and energy expenditure (EE) in male mice at 18 weeks of age across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (WT, n = 11-12; KO, n = 10-12). (H-J) Food intake, physical activity, and energy expenditure in female mice at 13 weeks of age (WT, n = 9; KO, n = 6). All data are presented as mean ± S.E.M.

Ctrp10-KO mice fed a low-fat diet have normal fasting-refeeding response and glucose homeostasis.

(A-B) Overnight fasted and refed blood glucose, serum insulin, triglyceride, cholesterol, non-esterified free fatty acids (NEFA), and β-hydroxybutyrate levels in male (A) and female (B) mice. (C-D) Blood glucose levels during glucose tolerance tests (GTT; C) and insulin tolerance tests (ITT; D) in WT (n = 17) and KO (n = 14) male mice at 12 weeks of age. (E-F) Blood glucose levels during glucose tolerance tests (GTT; E) and insulin tolerance tests (ITT; F) in WT (n = 9) and KO (n = 6) female mice at 20 and 21 weeks of age, respectively. All data are presented as mean ± S.E.M. * P < 0.05 (two-way ANOVA with Sidak’s post hoc tests).

CTRP10-deficient female mice on a low-fat diet develop obesity with age

Because female KO mice were slightly heavier at 20 weeks of age (Fig. 2C), we suspected the weight may diverge further with age. Consequently, we monitored the body weight of female mice fed LFD over an extended period. Indeed, the female KO mice gained significantly more weight and adiposity with age (Fig. 4A-C). Consistent with greater adiposity, the adipocyte cell size (cross-sectional area) was also significantly larger in both gonadal (visceral) white adipose tissue (gWAT) and inguinal (subcutaneous) white adipose tissue (iWAT) (Fig. 4D-E). By the time the mice reached 40 weeks of age, female KO mice weighed ∼ 6 g (20 %) heavier than the WT controls. Increased weight gain over time was not attributed to differences in food intake, as measured manually over a 24 h period (Fig. 4F). Fecal output, frequency, and energy content were also not different between genotypes (Fig. 4G), suggesting that weight gain was not due to greater nutrient absorption. Deep colon temperatures in both light and dark cycle were also not different between genotypes (Fig. 4H). Indirectly calorimetry analyses also revealed no significant differences between genotypes in food intake, physical activity, and energy expenditure across the circadian cycle and metabolic states (ad libitum fed, fasted, refed) (Fig. 4I-K). Despite significantly greater body weight and adiposity, female KO mice had the same metabolic profile as the lean WT controls. There were no differences in fasting blood glucose, serum insulin, triglyceride, cholesterol, NEFA, and β-hydroxybutyrate levels between genotypes (Fig. 4L). Interestingly, VLDL-TG levels were lower in female KO mice whereas HDL-cholesterol level was not different between genotypes (Fig. 4M). Direct assessments of glucose handling capacity and insulin sensitivity by glucose and insulin tolerance tests, respectively, also revealed no differences between genotypes (Fig. 4N-O). Together, these data indicate that Ctrp10-KO female mice fed LFD develop obesity, but preserve a largely healthy metabolic profile similar to the much leaner WT female mice.

Ctrp10-KO female mice on a low-fat diet develop obesity with age.

(A) Body weights over time of WT and KO female mice fed a low-fat diet (LFD). (B) Representative image of WT and KO female on LFD for 40 weeks. (C) Body composition analysis of WT (n = 9) and KO (n = 6) female mice fed a LFD. (D) Representative H&E stained histology of gonadal white adipose tissue (gWAT) and the quantification of adipocyte cell size (n = 6 per genotype). Scale bar = 100 μM. (E) Representative H&E stained histology of inguinal white adipose tissue (iWAT) and the quantification of adipocyte cell size (n = 6 per genotype). Scale bar = 100 μM. (F) 24-hr food intake data measured manually. (G) Fecal frequency, fecal weight, and fecal energy over a 24 hr period. (H) Deep colon temperature measured at the light and dark cycle. (I-K) Food intake, physical activity, and energy expenditure in female mice across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (WT, n = 9; KO, n = 6). Indirect calorimetry analysis was performed after female mice were on LFD for 30 weeks. (L) Overnight (16-hr) fasted blood glucose, serum insulin, triglyceride, cholesterol, non-esterified free fatty acids, and β-hydroxybutyrate levels. (M) Very-low density lipoprotein-triglyceride (VLDL-TG) and high-density lipoprotein-cholesterol (HDL-cholesterol) analysis by FPLC of pooled (n = 6-7 per genotype) mouse sera. (N) Blood glucose levels during glucose tolerance tests (GTT). (O) Blood glucose levels during insulin tolerance tests (ITT). GTT and ITT were performed when the female mice reached 28 and 29 weeks of age, respectively. WT, n = 9; KO, n = 6.

Rapid weight gain in CTRP10-deficient female mice fed a high-fat diet

Next, we challenged the mice with a HFD to determine if the sex-dependent effects on body weight become more pronounced. When fed a HFD, body weight gain and body composition were not different between genotypes in male mice (Fig. 5A-B). Food intake, physical activity, and energy expenditure were also not different between genotypes in male mice across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (Fig. 5C-E). In striking contrast, female KO mice gained weight rapidly on HFD (∼9 g or 28% heavier) and had greater adiposity than the WT controls (Fig. 5F-H). Surprisingly, food intake, physical activity, and energy expenditure were not significantly different between genotypes in female mice (Fig. 5I-K). The ANCOVA analysis of energy expenditure using body weight as a covariate also did not reveal any differences between genotypes in female mice (Fig. 5 L). Interestingly, the respiratory quotient (RER) was significantly lower in female KO mice relative to WT controls, especially during fasting and refeeding (Fig. 5M), suggesting a greater reliance on lipid substrates for energy metabolism during those periods. Together, these data indicate that CTRP10 is required for female-specific body weight control in response to caloric surplus, but neither food intake, physical activity level, nor energy expenditure could account for the marked increase in body weight and adiposity.

Sexually dimorphic response of Ctrp10-KO mice to an obesogenic diet.

(A) Body weights over time of WT and KO male mice fed a high-fat diet (HFD). (B) Body composition analysis of WT (n = 17) and KO (n = 14) male mice fed a HFD for 9 weeks. (C-E) Food intake, physical activity, and energy expenditure in male mice across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (WT, n = 11; KO, n = 11). Indirect calorimetry analysis was performed after male mice were on HFD for 10 weeks. (F) Body weights over time of WT and KO female mice fed a high-fat diet. (G) Representative image of WT and KO female mice after 13 weeks of high-fat feeding. (H) Body composition analysis of WT (n = 17) and KO (n = 13) female mice on HFD for 6 weeks. (I-K) Food intake, physical activity, and energy expenditure in female mice (WT, n = 11-12; KO, n = 12) across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed). Indirect calorimetry analysis was performed after female mice were on HFD for 6 weeks. (L) ANCOVA analysis of energy expenditure using body weight as a covariate. (M) Respiratory exchange ratio (RER). All data are presented as mean ± S.E.M. * P < 0.05; ** P < 0.01; *** P < 0.001; **** P < 0.0001

Obesity is uncoupled from insulin resistance and dyslipidemia in CTRP10-deficient female mice fed a HFD

We again measured fasting and refeeding responses in WT and Ctrp10-KO mice fed a HFD. No differences in fasting and refeeding blood glucose, serum insulin, triglyceride, cholesterol, NEFA, and β-hydroxybutyrate levels were noted between genotypes in male mice (Fig. 6A). Female KO mice, however, had higher fasting blood glucose and serum insulin levels, and lower β-hydroxybutyrate levels compared to the WT controls (Fig. 6B). In the refed state, serum insulin levels continued to be significantly higher in female KO mice. Unlike the WT female mice where refeeding markedly lowered serum β-hydroxybutyrate (ketone) levels as expected, KO female mice appeared unable to suppress serum β-hydroxybutyrate levels in response to refeeding (Fig. 6B).

Ctrp10-KO mice on a high-fat diet have normal glucose and insulin tolerance.

(A-B) Overnight fasted and refed blood glucose, serum insulin, triglyceride, cholesterol, non-esterified free fatty acids (NEFA), and β-hydroxybutyrate levels in male (A) and female (B) mice fed a HFD for 10 weeks. (C-D) Blood glucose levels during glucose tolerance tests (GTT; C) and insulin tolerance tests (ITT; D) in WT (n = 17) and KO (n = 14) male mice fed a HFD for 10 weeks. (E-F) Blood glucose levels during glucose tolerance tests (GTT; E) and insulin tolerance tests (ITT; F) in WT (n = 16) and KO (n = 12) female mice fed a HFD for 8 weeks. (G) VLDL-TG and HDL-cholesterol analysis by FPLC of pooled female mouse sera. All data are presented as mean ± S.E.M. ** P < 0.01; *** P < 0.001; **** P < 0.0001 (two-way ANOVA with Sidak’s post hoc tests for fasted/refed data).

Consistent with the fasting blood glucose and insulin data, glucose handling capacity and insulin sensitivity assessments by glucose and insulin tolerance in HFD-fed male mice also revealed no differences between genotypes (Fig. 6C-D). Female KO mice, however, had higher fasting blood glucose and serum insulin levels suggesting the presence of mild insulin resistance (Fig. 6B). We therefore expected to see differences in either glucose and/or insulin tolerance tests. To our surprise, the rate of glucose clearance in response to glucose or insulin injection was virtually identical between WT and KO female mice (Fig. 6E-F), suggesting no difference in insulin sensitivity between genotypes. VLDL-TG and HDL-cholesterol profiles were also indistinguishable between WT and KO female mice (Fig. 6G). Altogether, these data indicate that CTRP10 is not required for metabolic homeostasis in male mice challenged with a HFD. In female mice, however, loss of CTRP10 markedly promotes weight gain in the face of caloric surplus, but, paradoxically, the excess adiposity is largely uncoupled from obesity-linked insulin resistance and dysregulated glucose and lipid metabolism.

Obesity is uncoupled from adipose dysfunction and hepatic steatosis in Ctrp10-KO female mice fed a HFD

Consistent with greater fat mass in visceral (gonadal) fat depot of Ctrp10-KO female mice (Fig. 7A), histological analysis and quantification also indicated significantly larger adipocyte cell size (Fig. 7B). Although the subcutaneous (inguinal) fat pad weight was also significantly heavier in female KO mice (Fig. 7C), the adipocyte cell size was marginally bigger but not significant (Fig. 7D). A bigger fat pad with only marginally larger cell size suggests greater adipocyte hyperplasia in the subcutaneous depot. Increased adipogenesis in response to caloric surfeit is known to be associated with improved systemic metabolic profile (66).

Ctrp10-KO female mice fed a HFD do not develop adipose tissue dysfunction and fatty liver.

(A) Representative images of dissected gonadal white adipose tissue (gWAT) and the quantification of gWAT weight in WT (n = 15) and KO (n = 12) female mice fed a HFD for 14 weeks. (B) Representative H&E stained histological sections of gWAT and the quantification of adipocyte cell size (n = 7 per genotype). Scale bar = 100 μM. (C) Representative images of dissected inguinal white adipose tissue (iWAT) and the quantification of iWAT weight in WT (n = 15) and KO (n = 12) female mice. (D) Representative H&E stained histological sections of iWAT and the quantification of adipocyte cell size (n = 7 per genotype). Scale bar = 100 μM. (E) Expression of genes associated with inflammation, fibrosis, ER and oxidative stress in gWAT and iWAT of WT (n = 6) and KO (n = 6) female mice fed a HFD for 14 weeks. Gene expression data were obtained from RNA-seq. (F-G) Quantification of hydroxyproline (marker of fibrosis) and malondialdehyde (MDA; marker of oxidative stress) in gWAT and iWAT. WT, n = 15; KO, n = 11. (H) Representative images of dissected liver and the quantification of liver weight in WT (n = 15) and KO (n = 12) female mice. (I) Representative H&E stained histological sections of liver and the quantification of hepatic lipid content (% lipid area; n = 7 per genotype). Scale bar = 100 μM. (J) Hepatic expression of genes associated with inflammation, fibrosis, ER and oxidative stress, lipid synthesis, and lipid catabolism in WT and KO female mice. Gene expression data were obtained from RNA-seq. (K-L) Quantification of hydroxyproline (marker of fibrosis) and malondialdehyde (MDA; marker of oxidative stress) in liver. WT, n = 15; KO, n = 11. All data are presented as mean ± S.E.M. * P < 0.05; ** P < 0.01.

Obesity is known to be associated with low-grade inflammation (67), fibrosis (68), and ER and oxidative stress (69, 70). Despite marked differences in body weight and adiposity, the expression of genes associated with inflammation (except for Ccr2), fibrosis, oxidative stress in gWAT and iWAT were not significantly different between female KO mice and WT controls (Fig. 7E). The expression of some genes associated with ER stress (e.g., Ddit3/CHOP, Atf4, Xbp1) in iWAT were actually lower in female KO mice (Fig. 7E). Corroborating the gene expression data, quantification of hydroxyproline (marker of fibrosis) and malondialdehyde (marker of oxidative stress) revealed no significant differences between genotypes (Fig. 7F-G).

The liver weight of female KO mice was modestly increased (Fig. 7F), but when normalized to body weight it was not significantly different from WT controls (2.76 % in WT and 2.60% in KO, P = 0.24). Histological analysis and quantification revealed no differences in hepatic lipid content (% lipid area) between genotypes (Fig. 7G). Interestingly, although hepatic fat content was similar between genotypes, the expression of lipogenic genes (e.g., Fasn) was lower and fat catabolism genes (e.g., Cpt2, Ppara, Acadl, Acadm, Acad11, Acadvl) was higher in female KO mice (Fig. 7H). The expression of genes associated with inflammation, fibrosis, ER and oxidative stress in liver were not significantly different between genotypes (Fig. 7H). Consistent with the gene expression data, quantification of hydroxyproline (marker of fibrosis) in the liver revealed no significant difference between genotypes (Fig. 7K). The Ctrp10 KO female mice, however, had higher levels of malondialdehyde (a marker of oxidative stress) in the liver, suggesting a modest increase in oxidative stress (Fig. 7L). Altogether, these data indicate that obesity is largely uncoupled from inflammation, fibrosis, ER and oxidative stress in Ctrp10 KO female mice.

Transcriptomic and pathway changes associated with the metabolically healthy obesity phenotype in Ctrp10 KO female mice.

To define the specific mechanisms mediating the female-specific effects of Ctrp10 ablation on favorable metabolic outcomes, four major metabolic tissues (gWAT, iWAT, liver, skeletal muscle) from female WT and KO mice fed a HFD were subjected to RNA-sequencing. Comparison of differentially expressed genes (DEGs) via limma (71) showed robust changes across tissues, with the largest changes seen in the liver (Fig. 8A-D). In liver, gWAT, and muscle, we observed comparable numbers of DEGs that were up- and down-regulated, whereas more genes were transcriptionally suppressed in the iWAT of Ctrp10 KO female mice (Fig. 8E, top panel). While significant DEGs were identified in all 4 tissues, only limited overlap was observed between the DEGs in each tissue (Fig. 8E, bottom panel). Gene set enrichment analyses of the DEGs highlighted distinct and shared processes up- or down- regulated across the four tissues (Fig. 8F). Pathways and processes related to lipid metabolism and estrogen receptor were the top-ranked up-regulated enrichments across tissues (Fig. 8F, top panel), whereas processes related to blood clotting and lipoprotein metabolism were the top-ranked down-regulated enrichments (Fig. 8F, bottom panel).

Transcriptomic analysis of liver, adipose tissue, and skeletal muscle of female Ctrp10 KO mice fed a high-fat diet.

(A-D) Cropped volcano plot views of all differentially expressed genes (DEGs, Log2(Fold Change) >1 or <-1 with a p-value <0.05) of the liver, gonadal white adipose tissue (gWAT), inguinal WAT (iWAT), or skeletal muscle (gastrocnemius). (E) Overlap analysis of tissue DEGs showing (top panel) expression unique to gonadal white adipose tissue (gW), inguinal white adipose tissue (iW), liver (L), or skeletal muscle (M). Percent (%) represents percent DEGs unique to each tissue. Bottom panel show DEGs shared across multiple tissues, with all the shared DEGs listed. (F) Enrichr analysis (129) of biological pathways and processes significantly (p<0.01) affected across the CTRP10 deficient female mice. Top pathways and processes derived from Gene Ontology (GO), Reactome (R-HAS), WikiPathway human (WP), and mammalian phenotype (MP). All up- or down-regulated DEGs across all tissues were used for analysis. The tissues contributing to the highest ranked pathways and processes are specified. n = 6 KO and 6 WT for RNA-seq experiments.

Of the DEGs, we found significant changes across tissues in relevant classes of genes that encode proteins involved in gene expression (e.g., transcription factors), signaling (e.g., receptors), tissue crosstalk (e.g., secreted proteins), and metabolism (Fig. 9). Notably, the nuclear receptor, Nr1d1 (also known as Rev-Erbα), is the only gene consistently suppressed across all four tissues (liver, gWAT, iWAT, and muscle) of Ctrp10 KO female mice (Fig. 8E lower panel and Fig. 9). Interestingly, global deletion of Nr1d1 promotes lipogenesis, adipose tissue expansion, and obesity (72, 73). Although the whole-body and adipose-specific Nr1d1 KO mice fed with HFD become markedly obese, the obesity is not accompanied by insulin resistance, adipose tissue inflammation and fibrosis (73, 74). Like the Ctrp10 KO female mice, HFD-fed mice lacking Nr1d1 can maintain a relatively healthy metabolic profile despite being strikingly obese. Since only male mice were used in these previous studies, we do not know whether female mice lacking Nr1d1 would also exhibit similar insulin-sensitive obesity phenotype. In WT mice, Nr1d1 acts as a transcriptional repressor of metabolic genes whose expression are upregulated by high-fat feeding; loss of Nr1d1 is thought to result in the de-repression of these genes, leading to greater lipid synthesis and fat mass accrual in response to caloric excess (74). Thus, the suppression of Nr1d1 expression—mimicking Nr1d1 deficiency—across tissues in HFD-fed Ctrp10 KO female mice may contribute to benign fat mass expansion without the accompanying adipose tissue fibrosis, inflammation, and oxidative stress.

Loss of CTRP10 induces significant and wide-spread alterations in the expression of key transcription factors, secreted protein, membrane receptors, and metabolism-associated genes.

(A-D) Selected genes from the DEG list of each tissue organized based on gene type (genes encoding transcription factors, secreted proteins, receptors, and proteins involved in metabolism) and ranked from highest to lowest row z-score. N = 6 per genotype

Among the upregulated genes, Fgf21 and Fgf1 was significantly elevated in the liver of Ctrp10 KO female mice (Fig. 9A). FGF21 is an hepatokine known to improve systemic insulin sensitivity and to promote a favorable metabolic profile in diet-induced obese mice (75). Likewise, FGF1 has been shown to dampen hepatic glucose output by suppressing adipose lipolysis (76), improve systemic insulin sensitivity by reducing adipose inflammation (77), and alleviate hepatic steatosis, inflammation, and insulin resistance (78). Thus, upregulated expression of Fgf21 and Fgf1 in Ctrp10 KO female mice could contribute to the MHO phenotype. In addition, the upregulated hepatic expression of IL-22 receptor (Il22ra1) in Ctrp10 KO female mice (Fig. 9A) may confer protection against obesity-associated fatty liver, inflammation and fibrosis (79-81). Further, a marked increase in uncoupling protein 3 (Ucp3) and Krüppel-like factor 15 (Klf15) expression in the skeletal muscle (Fig. 9C) may promote lipid utilization and help mitigate lipid-induced insulin resistance in Ctrp10 KO female mice (82, 83). Taken together, these combined changes—at the level of gene expression and biological pathways and processes across tissues—acting in concert likely contribute to the apparently healthy obesity phenotype seen in the KO female mice.

Conservation of mouse DEG co-correlation in humans highlights sex-specific gene connectivity

Next, we asked whether the female-specific transcriptomic effects across tissues were conserved in humans. To address this, we analyzed transcriptional co-correlation of mouse DEG (Fig. 10) orthologues in GTEx (84), consisting of 210 males and 100 females filtered for comparison of gene expression across tissues (85, 86). Hierarchical clustering of transcriptional correlation of the orthologous DEGs among 4 metabolic tissues— subcutaneous and visceral white adipose tissue, liver, and skeletal muscle—showed differing patterns of gene connectivity between females (Fig. 10A) and males (Fig. 10B). When grouped according to sex in each tissue, the degree of sex-specific gene correlation pairs of DEGs orthologues showed the most significant differences in subcutaneous adipose tissue (Fig. 10C). Given the whole-body metabolic effects of Ctrp10 ablation in mice, we further examined the degree of sex-dependent DEG co-correlation across metabolic tissues. This analysis showed that human orthologue genes in subcutaneous adipose tissue (Fig. 10D, top row) and liver (Fig. 10D, third row) also exhibited highly significant sex differences in their transcriptional correlation with other DEGs across key metabolic tissues (Fig. 10D). These analyses highlight the sex-specificity of CTRP10 DEG orthologues in humans, suggest possible sex-biased mechanisms of tissue crosstalk, and overall underscores the conservation of the sex- dependent metabolic function of CTRP10.

GTEx genetic co-correlation of mouse differentially expressed gene (DEG) orthologues.

(A-B) Heatmaps showing biweight midcorrelation (bicor) coefficient among human tissue DEG orthologues in females (A) and males (B) in GTEx. Y-axis color indicates tissue of origin, P-value based on students’ regression P-value. (C) T-tests between correlation coefficient in males and females among all DEG orthologue gene pairs for subcutaneous (SubQ) adipose tissue, visceral (visc) adipose tissue, liver, and skeletal muscle. (D) the same as in C, except comparisons are shown for all gene-gene pairs between tissues. For example, the top left graph compared the connectivity of males (blue color) vs females (green color) for correlation between subcutaneous (SubQ) and visceral (Visc) adipose tissue DEG orthologues.

Discussion

Our current study has established a novel function for CTRP10 in modulating body weight in a sex-specific manner. When mice were fed a control LFD, female Ctrp10-KO mice developed obesity with age; increased adiposity, however, did not impair insulin action and glucose and lipid metabolism. When challenged with an obesogenic diet, female Ctrp10-KO mice gained weight rapidly. Despite having strikingly higher adiposity and weighing ∼10-11 g (∼28%) more, female KO mice fed a HFD exhibited a metabolic profile largely indistinguishable from the much leaner WT controls. Although female KO mice had higher fasting glucose and insulin levels, direct assessments of glucose metabolism and insulin sensitivity by glucose and insulin tolerance tests, however, revealed no differences between genotypes. Except having lower fasting ketone (β-hydroxybutyrate) levels, the fasting lipid profile, as well as VLDL-TG and HDL-cholesterol levels, of Ctrp10-KO female mice resembled the WT controls. The hepatic fat content was also comparable between genotypes. Global transcriptomic profiling across different fat depots and liver did not reveal gene expression signatures associated with elevated inflammation, fibrosis, and ER and oxidative stress. Altogether, these findings suggest that CTRP10 deficiency promotes obesity in females but it also uncouples obesity from insulin resistance, dyslipidemia, steatosis, inflammation, and oxidative stress. Thus, Ctrp10-KO female mice represent a novel model of female obesity with largely preserved insulin sensitivity and metabolic health.

Our findings help inform ongoing studies on metabolically healthy obese (MHO) humans (87-92). Because the criteria used to define MHO differs between studies, there is an ongoing debate regarding the prevalence of MHO and what fraction of the MHO population is insulin-sensitive and metabolically healthy (93, 94). Nevertheless, among the obese individuals, there clearly exists a subgroup that maintains long-term normal insulin sensitivity and does not appear to develop any component of the metabolic syndrome (93). MHO is observed in both sexes, but more common in females (6). The underlying mechanism(s) that uncouple obesity from adverse metabolic health in MHO is not well understood (7, 89, 93). It is currently unknown whether males and females with MHO use similar or distinct mechanism to maintain insulin sensitivity and metabolic health. Our findings in Ctrp10-KO female, but not male, mice suggest that there may be female-biased mechanism that prevents metabolic deterioration in the face of obesity, thus underscoring the utility of the Ctrp10-KO mice as a female mouse model of MHO.

Obesity is frequently associated with insulin resistance, dyslipidemia, fatty liver, oxidative stress, and chronic low-grade inflammation (67, 95, 96). The mechanisms that link obesity to metabolic dysfunctions are complex and multifactorial. There are limited number of mouse models described where obesity is uncoupled from insulin resistance and metabolic health (97-100); in some studies, however, only male mice were used or that the sex of the animals was not specified. In the case of aP2/FABP4 KO male mice, the uncoupling of obesity from insulin resistance was attributed to a marked decrease in TNF-α expression in adipose tissue (98). In the case of adiponectin overexpression in leptin-deficient (ob/ob) male and female mice, a dramatic expansion of the subcutaneous fat pad is thought to promote lipid sequestration in adipose compartment, thus preventing ectopic lipid deposition in non-adipose tissues (e.g., liver, pancreas, muscle) that would otherwise induce insulin resistance (97). Massive obesity with preserved insulin sensitivity is also observed in leptin-deficient (ob/ob) male and female mice overexpressing the mitochondrial membrane protein, mitoNEET (100). The benign obesity is attributed to the inhibition of iron transport into mitochondria by mitoNEET, leading to reduced mitochondrial activity, fatty acid oxidation, and oxidative stress (100). In the Brd2 hypomorphic mice, severe obesity with lower blood glucose and enhanced glucose tolerance is due to a combination of hyperinsulinemia and marked reduction in macrophage infiltration into fat depot (101). Lastly, in male mice fed a high starch diet, uncoupling of obesity from insulin resistance is associated with lower ceramide levels in liver and skeletal muscle (99). In all these cases, the uncoupling of obesity from metabolic dysfunction is seen in either male mice only (female mice were not included) or both sexes. These previous studies suggest that multiple mechanisms, not mutually exclusive, can contribute to the MHO phenotype in different mouse models.

In our study, loss of CTRP10 in female mice largely uncoupled obesity from insulin resistance, dyslipidemia, steatosis, inflammation, and ER and oxidative stress. The preservation of insulin sensitivity in Ctrp10 KO female mice is due, at least in part, to the absence of obesity-linked adipose and liver inflammation, fibrosis, and oxidative stress. These phenotypes associated with a favorable metabolic profile are also observed in MHO individuals (88, 102). A healthy adaptive remodeling of white adipose tissues in response to caloric surfeit helps preserve the storage and secretory function of adipocytes (103). The expansion of benign adipose tissues further serves to sequester circulating lipids and prevent their ectopic deposition in non-adipose tissue (e.g., liver and skeletal muscle) which can impair insulin action (104, 105). The MHO phenotype seen in Ctrp10 KO female mice reinforce the notion that adipose tissue health, rather than abundance, is an important determinant of metabolic health in obesity.

Because lipidomic analysis was not performed—a limitation of this study—we do not know whether Ctrp10-KO female mice have reduced ceramide or diacylglycerol levels in liver and skeletal muscle, two lipid species known to antagonize insulin action (106, 107). However, our global transcriptomic and pathway enrichment analysis across visceral and subcutaneous fat depots, liver, and skeletal muscle highlighted the relevant up- and down-regulated pathways and processes (e.g., lipid and lipoprotein metabolism, signaling) that may contribute to the MHO phenotype in Ctrp10-KO female mice. How these changes across tissues help to suppress the deleterious effects of obesity and maintain an apparently healthy metabolic profile in Ctrp10-KO female mice remains to be fully understood. Part of the mechanism may be attributable to the suppression of Nr1d1 and the upregulated expression of Fgf1, Fgf21, Il22ra1, Ucp3, Klf15. Altered expression of these genes are known to reduce obesity-linked inflammation, oxidative stress, steatosis, and insulin resistance.

In many single-gene KO mouse models where both sexes are examined, it is often the males that show a more pronounced metabolic phenotype. It is known that C57BL/6 female mice generally gain significantly less weight on HFD compared to male mice (108). Therefore, it is intriguing that Ctrp10-KO female mice became obese on a control LFD and gained weight rapidly when fed an obesogenic diet. After twelve weeks on HFD, the body weight of female KO mice was approaching that of WT male mice fed the same diet. What mechanism underlies the sexually dimorphic requirement of CTRP10 for body weight control? We know that the obesity phenotype was not attributed to differences in food intake, physical activity level, body temperature, and energy expenditure between WT and KO female mice. We assume that the methods used to quantify these physiologic parameters are sensitive enough to detect small differences that can give rise to divergent body weight over time. Quantification of fecal output and fecal energy content also revealed no differences between genotypes. Thus, loss of CTRP10 did not affect macronutrient intake and absorption. Although we cannot fully rule out the CNS function of CTRP10/C1QL2 (62), our data do not support a central role for CTRP10 in modulating food intake behavior, locomotor activity, and energy expenditure that affect body weight in female mice.

It is known that reduced estrogen level by ovariectomy or blocking estrogen action in estrogen receptor (ERα) KO mice will cause obesity and metabolic dysfunction in female mice fed a HFD (109-111). Conversely, estradiol supplementation decreases HFD-induced weight gain and improves glucose tolerance and insulin sensitivity (112-114). Estrogen also has the effect of reducing food intake, and promoting physical activity and energy expenditure (115-119). In our study, loss of CTRP10 promotes obesity without altering food intake, physical activity, and energy expenditure. While estrogen’s role cannot be completely ruled out, the fact that female Ctrp10-KO mice developed obesity with largely preserved metabolic health suggest that factors other than altered estrogen level contribute to the insulin-sensitive obesity phenotype. Future studies are warranted to uncover what factor(s) is causally contributing to obesity in female mice lacking CTRP10.

The sex-dependent effects of CTRP10 on metabolism and tissue transcriptomes appear to be conserved in humans. When the human orthologues of the mouse DEGs were used to interrogate the GTEx data, clear patterns of gene connectivity within and across metabolic tissues in females and males were observed, with the strongest sex-specific gene correlations seen in subcutaneous adipose tissue and liver. These findings provide further evidence that CTRP10 modulates tissue transcriptome in a sex-dependent manner. Further, our analyses of sex-dependent DEG co-correlation across metabolic tissues also suggest possible sex-biased mechanisms of inter-organ metabolic signaling between adipose tissue and liver.

CTRP10 has been previously shown to bind to the adhesion GPCR, brain angiogenesis inhibitor-3 (Bai3/Adgrb3) (120). Bai3 is expressed in the brain and peripheral tissues, and it is a promiscuous GPCR that can bind to multiple ligands. In addition to CTRP10 (C1QL2), Bai3 also binds to CTRP11 (C1QL4), CTRP13 (C1QL3), CTRP14 (C1QL1), neuronal pentraxins, and reticulon 4 (RTN4) receptor (56, 58, 59, 120-122). A constitutive, whole-body KO of Bai3 mouse models have recently been generated (123, 124). Both male and female Bai3 KO mice fed a standard chow have significantly lower body weight, beginning at weaning (3 weeks old) and continue into adulthood (123, 124). Lower body weight in Bai3 KO mice of either sex is attributed to a reduction in both lean and fat mass, and is associated with higher energy expenditure and reduced food intake in male mice (123). The impact of Bai3 deficiency on systemic metabolism in response to a high-fat diet was not examined. The Ctrp10 KO mice do not phenocopy the phenotypes of the Bai3 KO mice. When fed a control low-fat diet, the body weight, food intake, and energy expenditure of Ctrp10 KO male mice were indistinguishable from WT controls. In striking contrast to Bai3 KO mice, female Ctrp10 KO mice fed a low-fat diet began to gain more weight around 20 weeks of age, and by 40 weeks had become visibly obese. While we did not rule out CTRP10-Bai3 signaling axis in modulating energy metabolism in peripheral tissues, our findings in Ctrp10 KO mice suggest that future works are needed to establish the molecular mechanisms that mediate the systemic metabolic function of CTRP10.

Several limitations of our current study are noted. We use a constitutive whole-body KO mouse model of CTRP10 to interrogate its function. It is unknown whether CTRP10 has a role during development that may influence sex-dependent postnatal weight gain with age or in response to a high-caloric diet. Future studies using conditional KO of Ctrp10 gene in adult mice can help address this issue. Although the lack of differences in food intake, physical activity, and energy expenditure between genotypes do not support a central role of CTRP10 in mediating the metabolic phenotypes of Ctrp10-KO female mice, a brain-specific KO of Ctrp10 gene is needed to definitively rule this out. Our phenotypic analyses in the context of metabolism are relatively comprehensive but not exhaustive. Although the metabolic profile of obese Ctrp10-KO female mice was largely indistinguishable from the much leaner WT controls, we do not know if some related aspect of metabolic health (e.g., blood pressure and heart function) may be altered in the absence of CTRP10 which we did not examine.

In summary, we have established the physiologic role and requirement of CTRP10 in modulating body weight in a female-specific manner. Importantly, loss of CTRP10 largely uncouples obesity from insulin resistance and metabolic dysfunction. The CTRP10-deficient female mice represent a unique and valuable model to help dissect female-biased mechanisms that help preserve metabolic health in the face of positive energy balance and increased adiposity.

Materials and methods

Mouse models

Eight-week-old mouse tissues (gonadal and inguinal white adipose tissues, interscapular brown adipose tissue, liver, heart, skeletal muscle, kidney, pancreas, cerebellum, cortex, hippocampus, hindbrain, and hypothalamus) from C57BL/6J male mice (The Jackson Laboratory, Bar Harbor, ME) were collected from fasted and refed experiments as we have previously described (125). For the fasted group, food was removed for 16 h (beginning 10 h into the light cycle), and mice were euthanized 2 h into the light cycle. For the refed group, mice were fasted for 16 h and refed with chow pellets for 2 h before being euthanized. Tissues (white and brown adipose tissues, liver, whole brain, kidney, spleen, heart, skeletal muscle, pancreas, small intestine, and colon) from C57BL/6J male mice fed a low-fat diet (LFD) or a high-fat diet (HFD) for 12 weeks were also collected as we have previously described (125).

The Ctrp10/C1ql2-KO mice (C57BL/6NCrl-C1ql2em1(IMPC)Mbp/Mmucd; stock number 050587-UCD) were generated using the CRISPR-cas9 method at UC Davis. The two guide RNAs (gRNA) used were 5’-CCGGCGCC GCTCCACCATTACCT-3’ and 5’-TCAGGCCACCCCATCCCCATCGG-3’. The Ctrp10 gene consists of 2 exons. The entire protein coding region spanning exon 1 and 2 was deleted. This KO strategy ensures a complete null allele for Ctrp10. The KO mice were maintained on a C57BL6/6J genetic background. Genotyping primers for WT allele were forward (m10-Com-F) 5’-TGTCGGGCTCTTCGACTCTCCA-3’ and reverse (m10-WT-R) 5’-GCATCTCGT AGTGAGCCGCTCC-3’. The size of the WT band was 360 bp. Genotyping primers for the Ctrp10 KO allele were forward (m10-Com-F) 5’-TGTCGGGCTCTTCGACTCTCCA-3’ and reverse (m10-Mut-R1) 5’-GTCCAATCAGCT TTCTCAAGTCTGG-3’. The size of the KO band was 422 bp. The genotyping PCR parameters were as follows: 94°C for 5 min, followed by 10 cycles of (94°C for 10 sec, 65°C for 15 sec, 72°C for 30 sec), then 25 cycles of (94°C for 10 sec, 55°C for 15 sec, 72°C for 30 sec), and lastly 72°C for 5 min. Due to the presence of GC rich sequences, 7% DMSO was included in the PCR genotyping reaction. Mice were generated by intercrossing Ctrp10 heterozygous (+/-) mice, supplemented with intercrossing WT or KO mice. Ctrp10 KO (-/-) and WT (+/+) controls were housed in polycarbonate cages on a 12-h light–dark photocycle with ad libitum access to water and food. Mice were fed either a control low-fat diet (LFD; 10% kcal derived from fat; # D12450B; Research Diets, New Brunswick, NJ) or a high-fat diet (HFD; 60% kcal derived from fat; #D12492, Research Diets). LFD was provided for the duration of the study, beginning at 5 weeks of age; HFD was provided for 14 weeks, beginning at 6-7 weeks of age. At termination of study, all mice were fasted for 2 h and euthanized. Tissues were collected, snap-frozen in liquid nitrogen, and kept at -80°C until analysis. All mouse protocols (protocol # MO22M367) were approved by the Institutional Animal Care and Use Committee of the Johns Hopkins University School of Medicine. All animal experiments were conducted in accordance with the National Institute of Health guidelines and followed the standards established by the Animal Welfare Acts.

Body composition analysis

Body composition analyses for total fat, lean mass, and water content were determined using a quantitative magnetic resonance instrument (Echo-MRI-100, Echo Medical Systems, Waco, TX) at the Mouse Phenotyping Core facility at Johns Hopkins University School of Medicine.

Indirect calorimetry

LFD- or HFD-fed WT and Ctrp10 KO male and female mice were used for simultaneous assessments of daily body weight change, food intake (corrected for spillage), physical activity, and whole-body metabolic profile in an open flow indirect calorimeter (Comprehensive Laboratory Animal Monitoring System, CLAMS; Columbus Instruments, Columbus, OH) as previously described (25). In brief, data were collected for three days to confirm mice were acclimatized to the calorimetry chambers (indicated by stable body weights, food intakes, and diurnal metabolic patterns), and data were analyzed from the fourth day. Rates of oxygen consumption (V˙ O2; mL·kg-1·h-1) and carbon dioxide production (V˙ CO2; mL·kg-1·h-1) in each chamber were measured every 24 min throughout the studies. Respiratory exchange ratio (RER = V˙ CO2/V˙ O2) was calculated by CLAMS software (version 5.66) to estimate relative oxidation of carbohydrates (RER = 1.0) versus fats (RER = 0.7), not accounting for protein oxidation. Energy expenditure (EE) was calculated as EE= V˙ O2× [3.815 + (1.232 × RER)] and normalized to lean mass. Because normalizing to lean mass can potentially lead to overestimation of EE, we also performed ANCOVA analysis on EE using body weight as a covariate (126). Physical activities were measured by infrared beam breaks in the metabolic chamber.

Measurements of 24 h food intake

To independently confirm the food intake data collected in the metabolic cage (CLAMS), we also performed 24 h food intake measurements manually. All mice were singly housed, with wire mesh flooring inserts over a piece of cage paper on the bottom of the cage. A known weight of food pellets was given to each mouse. Twenty-four hours later, the leftover food pellets remaining on the flooring insert, along with any spilled crumbs on the cage paper, were collected and weighed. Thus, food intake was corrected for spillage.

Glucose, insulin, pyruvate, and lipid tolerance tests

All tolerance tests were conducted as previously described (21, 24, 29). For glucose tolerance tests (GTTs), mice were fasted for 6 h before glucose injection. Glucose (Sigma, St. Louis, MO) was reconstituted in saline (0.9 g NaCl/L), sterile-filtered, and injected intraperitoneally (i.p.) at 1 mg/g body weight (i.e., 10 μL/g body weight). Blood glucose was measured at 0, 15, 30, 60, and 120 min after glucose injection using a glucometer (NovaMax Plus, Billerica, MA). For insulin tolerance tests (ITTs), food was removed 2 h before insulin injection. 6.5 μL of insulin stock (4 mg/mL; Gibco) was diluted in 10 mL of saline, sterile-filtered, and injected i.p. at 0.75 U/kg body weight (i.e., 10 μL/g body weight). Blood glucose was measured at 0, 15, 30, 60, and 90 min after insulin injection using a glucometer (NovaMax Plus).

Fasting-Refeeding insulin tests

Mice were fasted overnight (∼16 h) then reintroduced to food as described (26). Blood glucose was monitored at the 16 h fast time point (time = 0 h refed) and at 1 and 2 hours into the refeeding process. Serum was collected at the 16 h fast and 2 h refed time points for insulin ELISA, as well as for the quantification of triglyceride, cholesterol, non-esterified free fatty acids (NEFA), and β-hydroxybutyrate levels.

Blood and tissue chemistry analysis

Tail vein blood samples were allowed to clot on ice and then centrifuged for 10 min at 10,000 x g. Serum samples were stored at -80°C until analyzed. Serum triglycerides (TG) and cholesterol levels were measured according to manufacturer’s instructions using an Infinity kit (Thermo Fisher Scientific, Middletown, VA). Non-esterified free fatty acids (NEFA) were measured using a Wako kit (Wako Chemicals, Richmond, VA). Serum β-hydroxybutyrate (ketone) concentrations were measured with a StanBio Liquicolor kit (StanBio Laboratory, Boerne, TX). Serum insulin levels were measured by ELISA according to manufacturer’s instructions (Crystal Chem, Elk Grove Village, IL; cat # 90080). Hydroxyproline assay (Sigma Aldrich, MAK008) was used to quantify total collagen content in liver and adipose tissues according to the manufacturer’s instructions. Lipid peroxidation levels (marker of oxidative stress) in the liver and adipose tissues were assessed by the quantification of malondialdehyde (MDA) via Thiobarbituric Acid Reactive Substances (TBARS) assay (Cayman Chemical, 700870) according to the manufacturer’s instructions.

Serum lipoprotein triglyceride and cholesterol analysis by FPLC

Food was removed for 2-4 hr (in the light cycle) prior to blood collection. Sera collected from mice were pooled (n = 6-7/genotype) and sent to the Mouse Metabolism Core at Baylor College of Medicine for analysis. Serum samples were first fractionated by fast protein liquid chromatography (FPLC). A total of 45 fractions were collected, and TG and cholesterol in each fraction was quantified.

Histology and quantification

Inguinal (subcutaneous) white adipose tissue (iWAT), gonadal (visceral) white adipose tissue (gWAT), and liver were dissected and fixed in formalin. Paraffin embedding, tissue sectioning, and staining with hematoxylin and eosin were performed at the Pathology Core facility at Johns Hopkins University School of Medicine. Images were captured with a Keyence BZ-X700 All-in-One fluorescence microscope (Keyence Corp., Itasca, IL). Adipocyte (gWAT and iWAT) cross-sectional area (CSA), as well as the total area covered by lipid droplets in hepatocytes were measured on hematoxylin and eosin-stained slides using ImageJ software (127). For CSA measurements, all cells in one field of view at 100X magnification per tissue section per mouse were analyzed. Image capturing and quantifications were carried out blinded to genotype.

Fecal bomb calorimetry and assessment of fecal parameters

Fecal pellet frequency and average fecal pellet weight were monitored by housing each mouse singly in clean cages with a wire mesh sitting on top of a cutout cardboard that lay at the bottom of the cage for fecal collection. The number of fecal pellets and their total weight was recorded at the end of 24 h period. Additional fecal pellets collected for 3 full days were combined and shipped to the University of Michigan Animal Phenotyping Core for fecal bomb calorimetry. Briefly, fecal samples were dried overnight at 50°C prior to weighing and grinding them to powder. Each sample was mixed with wheat flour (90% wheat flour, 10% sample) and formed into 1.0 g pellet, which was then secured into the firing platform and surrounded by 100% oxygen. The bomb was lowered into a water reservoir and ignited to release heat into the surrounding water. These data were used to calculate fecal pellet frequency (bowel movements/day), average fecal pellet weight (g/bowel movement), fecal energy (cal/g feces), and total fecal energy (kcal/day).

Tissue library preparation and RNA sequencing

Total RNA was isolated from tissues using Trizol reagent (Thermo Fisher Scientific) according to the manufacturer’s instructions. Library preparation and bulk RNA sequencing of liver, skeletal muscle (gastrocnemius), gonadal white adipose tissue (gWAT), and inguinal white adipose tissue (iWAT) of HFD-fed Ctrp10-KO female mice and WT controls were performed by Novogene (Sacramento, California, USA) on an Illumina platform (NovaSeq 6000) and pair-end reads were generated. Sample size: 6 WT and 6 KO for each tissue. All raw sequencing files are available from the NIH Sequence Read Archive (SRA) accession PRJNA971939.

Mouse RNA-Sequencing analysis

Transcript features were assembled from raw fastq files and aligned to the current version of mouse transcriptome (Mus_musculus.GRCm39.cdna) using kallisto -aln (128). Version-specific Ensembl transcript IDs were linked to gene symbols using biomart. Estimated counts were log normalized and filtered for a limit of sum >5 across all samples. Logistic regressions comparing WT vs KO samples across tissues were performed using limma (71). Differential expression results were visualized using available R packages in CRAN: ggplot2, ggVennDiagram and pheatmap. Gene set enrichment analyses of the DEGs were performed using Enrichr (129). Scripts for analyses and visualization are available at https://github.com/Leandromvelez/CTRP10-Manuscript-DEG-Sex-specific-connectivities-and-integration/

Human sex difference analysis

All the datasets and scripts to perform analyses are available at: https://github.com/Leandromvelez/CTRP10-Manuscript-DEG-Sex-specific-connectivities-and-integration/. Male and Female human data were obtained from Genotype-Tissue Expression (GTEx) (84) and filtered for sufficient comparison of inter-tissue transcript correlation as described (85, 86). CTRP10/C1QL2 and other human orthologues for mouse differentially expressed genes (DEGs) were identified by intersecting mouse gene symbols with known human orthologues from the vertebrate homology resource at Mouse Genome Informatics (MGI) (130). Co-correlation between all human orthologues DEGs were calculated in either self-reported male or female subjects in GTEx using the bicorAndPvalue() function in Weighted Genetic Coexpression Network Analysis (WGCNA) package (131). To compare sex-differences of regression coefficients, wilcoxon t-tests were compared between coefficients using the R package ggpubr.

Quantitative real-time PCR

Total RNA was isolated from tissues using Trizol reagent (Thermo Fisher Scientific). Purified RNA was reverse transcribed using an iScript cDNA Synthesis Kit (Bio-rad). Real-time quantitative PCR analysis was performed on a CFX Connect Real-Time System (Bio-rad) using iTaqTM Universal SYBR Green Supermix (Bio-rad) according to manufacturer’s instructions. Data were normalized to the stable housekeeping gene β-actin or 36B4 (encoding the acidic ribosomal phosphoprotein P0) and expressed as relative mRNA levels using the ΔΔCt method (132). Real-time qPCR primers used to assess Ctrp10 expression across mouse tissues were: Ctrp10 forward, 5’-CGGCTTCATGAC ACTTCCTGA-3’ and reverse, 5’-AGCAGGGATGTGTCTTTTCCA-3’. qPCR primers used to confirm the absence of Ctrp10 in KO mice were: forward (qPCR-m10-F2), 5’-CACGTACCACATTCTCATGCG-3’ and reverse (qPCR-m10-R1), 5’-TCGTAATTCTGGTCCGCGTC-3’.

Statistical analyses

Sample size is indicated in figure and/or figure legend. All results are expressed as mean ± standard error of the mean (SEM). Statistical analysis was performed with Prism 9 software (GraphPad Software, San Diego, CA). Data were analyzed with two-tailed Student’s t-tests, one-way ANOVA or two-way ANOVA (with Sidak’s post hoc tests). 2-way ANOVA was used for body weight over time, fasting-refeeding response, and all tolerance tests. P < 0.05 was considered statistically significant.

eLife Assessment

This manuscript presents a detailed characterization of male and female wild-type and CTRP10 knockout mice, revealing that knockout mice develop female-specific obesity that is largely uncoupled from metabolic dysfunction. The data are convincing, and the work is a valuable contribution to understanding how obesity is coupled to metabolic dysfunction, and how this can occur in a sex-specific manner.

Reviewer #1 (Public Review):

Summary:

The manuscript by Chen et al. presents a detailed metabolic characterization of male and female WT and CTRP10 knockout mice. The main finding is that female KO mice become obese on both low-fat and high-fat diets but without evidence of marked insulin resistance, hepatic steatosis, dyslipidemia, or increased inflammatory markers. The authors performed a detailed transcriptomic analysis and identified differentially expressed genes that distinguish high-fat diet-fed CTRP10 KO from WT control mice. They further show that this set of genes exhibits cross-correlation in human tissues, and that this is greater in females than in males. The data indicate that the CTRP10 KO model may be useful to understand how obesity and metabolic dysfunction are coupled to each other, and how this occurs by a sex-biased mechanism.

Strengths:

The work presents a large amount of data, which has been carefully acquired and is convincing. The transcriptomic analysis will further help to define what pathways are associated with obesity, but not necessarily with metabolic dysfunction. The manuscript will be of interest to investigators studying metabolic diseases, and to those studying sex-specific differences in metabolic physiology. The limitations of the study are acknowledged, including that a whole-body knockout was used. The cause of the increased body weight is not entirely clear, despite the careful and detailed analysis that was performed. Notwithstanding these limitations, the phenotype is interesting, and this work will establish a basis for further work to understand the mechanisms that are involved.

Weaknesses:

Genes identified as DEGs in the mouse RNAseq data set were used to identify a set of human orthologous transcripts and the abundances of these transcripts were correlated with each other in Figure 10. This identified a greater correlation ("connectivity") in subQ adipose compared to other tissues, and in females compared to males. The description of how this analysis was done could be clearer. In some cases, the text refers to the software that was used without describing the goal of the analysis. In other instances, specialized terminology was used (e.g. "biweight midcorrelation") without defining what this means.

Reviewer #2 (Public Review):

Summary:

In the current study, the authors investigated the role of loss of CTRP10 results in female obesity with preserved metabolic health. The overall conclusion is supported by the experimental data that CTRP10 negatively regulates body weight in females and that loss of CTRP10 results in benign obesity with largely preserved insulin sensitivity and metabolic health. The authors have shown the role of sex differences in the metabolically healthy obese (MHO) phenotype, which may increase the scope for research in this area.

Strengths:

The study provides a detailed idea of how genes are regulated in a sex-dependent manner.

Weaknesses:

Mechanistic details are missing.

Reviewer #3 (Public Review):

Summary:

This study examines the impact of CTRP10/C1QL2 absence on obesity and metabolic health in mice. Female mice lacking CTRP10 tend to develop obesity, particularly on a high-fat diet. Surprisingly, they do not display the typical metabolic traits associated with obesity, like fatty liver or glucose intolerance. This indicates a disconnection between weight gain and metabolic issues in these female mice. The research underscores the need to understand sex-specific factors in how obesity influences metabolic health.

Strengths:

The study provides compelling evidence regarding Ctrp10's role in female-specific metabolic regulation in mice, shedding light on its potential significance in metabolically healthy obese (MHO) individuals.

Weaknesses:

-The analysis and description of sex-specific human data require more details to highlight the relevance of Ctrp10 mouse data and the analysis of differentially expressed genes in humans.

-There's a lack of analysis regarding secreted Ctrp10 under various dietary conditions.

-The study didn't assess adipose tissue function to evaluate metabolic health.

Author response:

The following is the authors’ response to the original reviews.

Reviewer #1 (Recommendations For The Authors):

Although the scripts are available at the github link that is shown, the Readme file is not available as a text file. Spreadsheets summarizing the RNA-seq data ought to be available for download, but these are not present. Likewise, are spreadsheets available for the data used to generate the plots in Fig. 10, so that the identities of particular, correlated genes can be viewed?

We have now included the excel sheet with all the DEGs shown in Figure 8-9 (Figure 8 – Source data 1-8). The source data include DEGs that are up- and down-regulated in gWAT, iWAT, liver, and skeletal muscle. The source data files (excel) are the standard output format. We have also updated the github (https://github.com/Leandromvelez/CTRP10-Manuscript-DEG-Sex-specific-connectivities-and-integration) to include a README file and updated the R scripts to annotate steps and processing considerations. In addition, the README file now contains drive links to the files used the unfiltered kallisto TPM and counts at the transcript-level, as well as resulting Differential Expression results based on genotype. Obviously, all criteria from aligned transcripts such as gene filtering and normalization are included in the scripts provided.

Several items would strengthen the work:

(1) Is a CTRP10 antibody available, and does the protein abundance correlate with the mRNA abundances that were assessed in Fig. 1?

Unfortunately, no validated antibody currently exists for CTRP10. Consequently, we were not able to assess protein abundance of CTRP10 in our study.

(2) Were there compensatory changes in the abundance of other CTRP family members? This might be observed at the protein, but not mRNA, level. It might be reasonable to test for the effects of liver, gWAT, skeletal muscle, and iWAT.

We observed no compensatory changes in other CTRP family members based on our RNA-seq data. Unfortunately, we do not have protein data for other CTRP family members.

(3) The gene expression changes shown in Fig. 9 are ranked according to z-score, but it is not clear how this is calculated. It would be helpful to indicate the log2 change in each case.

The z-score is a very commonly used method to show DEGs in studies involving RNA-seq data. We calculate the z-score based on the gene transcript source data (Fig. 8 – Source data 1-8). Z-score is defined as z = (x-μ)/σ, where x is the raw score (gene transcript level), μ is the population mean (mean of gene expression across both WT and KO samples), and σ is the population standard deviation. In essence, the z-score is the raw score minus the population mean, divided by the population standard deviation. We now included this information in Fig. 9 legend.

(4) In Fig. 6, female HFD-fed KO mice had increased glucose (and insulin) after an overnight fast, but increased glucose was not observed in the GTT data. Possibly, this is because the mice were fasted for only 6h for the GTT. This might be mentioned during the description of these data, on lines 221-224. However, this also raises the question of whether there is a difference in the rate of gluconeogenesis (or possibly glycogenolysis for the 6h data) in the KO compared to the controls. Understanding this would require the use of tracers, and is reasonably beyond the scope of this study, but might be mentioned in the discussion.

Per reviewer’s suggestion, we have included this in the “limitation section” of the discussion.

Reduced RER in the HFD-fed female mice might begin to suggest a mechanism since this suggests the mice might have decreased oxidation of carbohydrates and increased oxidation of fat compared to control animals. A glucose tracer might be used to test whether more glucose is stored and, if so, in what tissue this occurs. Possibly, this could be done ex vivo on isolated tissues or cells. Again, this is reasonably beyond the scope of the present study.

Per reviewer’s suggestion, we have included this in the “limitation section” of the discussion.

(5) The discussion includes a brief discussion of the role of estrogen and suggests that in CTRP10 KO mice there are differences in other factors that would be needed to explain the phenotype. Although it is agreed that this is likely the case, estrogen levels were not measured in the present study. It seems like this would be important to study, and might shed light on the female-specific phenotype.

We have now included serum estrogen data. No significant differences in estrogen levels were seen between WT and KO female mice fed either a low-fat diet (Fig. 4 – figure supplement 1) or a high-fat diet (Fig. 5 – figure supplement 2).

Reviewer #2 (Recommendations For The Authors):

While the concept is potentially exciting, there are major problems with the current manuscript. It lacks the mechanistic details behind MHO.

(1) There is a significant gap that was not addressed by the authors. How exactly does CTRP10 lead to the activation of proteins like Fgf1, Fgf21, Il22ra1, Ucp3, and Klf15 in Ctrp10 knockout female mice? Is it likely that CTRP10 regulates these proteins via indirect mechanisms?

We acknowledge that the lack of mechanistic understanding of how CTRP10 loss-of-function leads to changes in gene expression is a major limitation of the study. We have highlighted this limitation in the discussion section.

• The author notes that Ctrp10 knockout female mice, particularly those on a high-fat diet lack Nr1d1 and can sustain a relatively healthy metabolic state. This is supported by the demonstrated upregulation of Fgf1, Fgf21, Il22ra1, Ucp3, and Klf15 in Ctrp10 knockout female mice. However, the mechanisms through which Ctrp10 knockout influences the expression of these molecules are not elucidated.

We acknowledge that this is a major limitation of the study. We have highlighted this limitation in the discussion section.

• How do you substantiate the role of age and a high-nutrient diet in the development of obesity in knockout female mice? However, it is still unclear whether administering a high-fat diet in >20 week age of mice can develop insulin resistance where obesity is developing in LFD.

When fed a low-fat diet, Ctrp10-KO female mice developed obesity with age and yet show little if any glucose intolerance or insulin resistance based on our glucose tolerance and insulin tolerance tests. For the HFD group, we are only comparing WT and KO mice on the same diet (not across diet). While WT mice on HFD gained significant amount of weight over time as expected, Ctrp10-KO female mice gain substantially higher amount of weight relative to WT littermates. Despite this, we did not observe a worsening of glucose tolerance and insulin resistance (based on GTT and ITT) in the KO female mice relative to WT controls as we would expect, since greater adiposity in HFD-fed mice generally correlated with worse metabolic outcomes.

(2) The authors should add the NR1D1 dependency study in female mice if possible.

To address would require the generation of Ctrp10/Nr1d1 double KO mouse model and to carry out the entire study again in these double KO mice. Although this suggestion by the reviewer is a good one, this is beyond the scope of the present study.

(3) NR1D1 represses the set of genes that promotes lipogenesis (the author should add some data that validates this statement).

The role of NR1D1 in regulating metabolic genes are extensively documented in the published literature. NR1D1 (also known as REV-ERBα) is a constitutive transcriptional repressor (PMID: 26044300; PMID: 27445394). Many metabolic genes that are normally represses by NR1D1 is de-repressed in mice lacking NR1D1 globally or in the tissue-specific manner (PMID: 26044300; PMID: 34350828; PMID: 22562834). Among the many NR1D1 target genes involved in lipid metabolism include: CD36, Plin2, Elovl5, Acss3 (from: PMID: 26044300); as well as Scd1, Scd2, Pnpla5, Acsl1, Fasn, Hadhb, and Oxsm (from: PMID: 34350828). We have included this information in the discussion section.

(4) The authors should study the effect of Ctrp10 overexpression in HFD-fed female mice and also with KO of CTRP10 in adult mice if possible.

The suggestion by the reviewer is a good one. However, this is beyond the scope of the study. We do not have a Ctrp10 conditional KO mouse model; as such, we could not study the effect of knocking out CTRP10 in adult mice. Overexpression studies are often considered non-physiological these days since the level of the overexpressed protein is generally much higher than the normal physiological level. For this reason, we did not attempt any overexpression study.

Reviewer #3 (Recommendations For The Authors):

Line 114: Could you please provide definitions for "GluK2" and "GluK4" for readers unfamiliar with these terms?

We have now provided definition for these terms.

Line 140: It's stated that skeletal muscle and the pancreas express similar levels of Ctrp10 as the brain. Please double-check and clarify this assertion for accuracy.

In mice, based on our own data (Fig. 1B), Ctrp10 expression in skeletal muscle and pancreas is comparable to that in the whole brain. In human, based on publicly available data (e.g., Genotype-Tissue Expression portal; GTex), brain expresses much higher level of CTRP10 transcript relative to other peripheral tissues.

Line 141: Have you investigated whether Ctrp10 levels in plasma change after refeeding? If not, consider exploring this aspect to enhance the comprehensiveness of the study.

No validated antibody currently exists for CTRP10. As such, we could not assess plasma level of CTRP10 after refeeding. We have included this as limitation of our study in the discussion section.

Lines 143-144: Clarify the age bracket of the animals used in the study. Additionally, have you observed similar responses, such as downregulation of Ctrp10 in response to refeeding, in both old and young mice in peripheral tissues?

We have now included the age of the mice (~10 weeks old) for the fasting refeeding study as shown in Fig. 1C in the result and method sections.

Lines 135-149: To complement the experiments shown in Fig 1B-D, provide data pertaining to females.

Ideally, we would like to have this data as well. However, to do this for females would involve 47 mice and the collection of 120 tissues (Fig. 1B; n = 10 per tissue), 390 tissues (Fig. 1C; n = 7-8 per tissue per fast or refed state), and 528 tissues (Fig. 1D; n = 11 per tissue per HFD or LFD). This would be a total of 1038 tissue samples. The main purpose of Fig. 1B-D is to demonstrate that Ctrp10 transcript is widely expressed and that its expression is modulated by nutritional (HFD vs. LFD) and metabolic (fast vs. refeed) states. These data provided a rationale to examine the metabolic phenotype in mice lacking CTRP10.

To address the reviewer’s point, we looked at the expression levels of CTRP10/C1QL1 between males and females in the Genotype-Tissue Expression (GTEx) database portal and it does not appear that there are sex differences in CTRP10 expression patterns in normal tissues.

Line 152: Can you provide evidence supporting the hypothesis that Ctrp10 is secreted into the circulation?

CTRP10 has a classic signal peptide sequence and the protein is secreted when expressed in HEK 293 cells (PMID: 18783346). We have shown previously that CTRP10 can be found in the FPLC-fractionated mouse serum using a polyclonal rabbit anti-mouse CTRP10 antibody we generated (PMID: 18783346); this antibody, however, does not work on tissue lysates (many non-specific bands). There is evidence in published literature to show that CTRP10/C1QL2 is clearly found circulating in human plasma. Some of the studies include: 1) Human C1QL2/CTRP10 is detected in the human plasma from UK BioBank (PMID: 37794186; C1QL2 is highlighted in page 335) and serum samples from pregnant females (PMID: 39062451; C1QL2 is highlighted in Table 2). We have included this information in the Introduction section.

Line 178: In Fig 4 D and E (and other figures in the paper), it would be more accurate to express adipocyte size in "μm²" instead of "uM2."

We have double checked and fixed this issue in the figure 4 and 7.

Line 259: Please specify the age of the animals used in the study.

In the method section, we did mention that LFD was provided for the duration of the study, beginning at 5 weeks of age; and that HFD was provided for 14 weeks, beginning at 6-7 weeks of age. Also, in Figure 2A and Figure 4A, the age of the mice is also indicated.

Lines 275-283 and 288-296: It would be more appropriate to move this content to the Discussion section for better contextualization.

We feel that the published information on NR1D1 and FGF21 should be mentioned in the result section so that the readers can immediately appreciate the significance of our data shown in Fig. 8 and 9. However, we also included similar information concerning NR1D1 in the discussion section for better contextualization as suggested.

Line 301: The section on DEG analysis requires additional details. How was the DEG analysis conducted? Were the DEGs from "wild type and KO mice" compared with "human DEGs regulated by sex"? Also, details about the phenotype of the human subjects and their association with obesity should be included. Additionally, discuss specific genes identified by the analysis and their relevance to the Ctrp10 story and human sex-specific gene connectivity analysis.

We have updated the section on DEG analysis and, related to reviewer comments above, significantly expanded the github repository, detailing an analytical walkthrough of all computational analyses performed. To clarify the human integration analysis, we have added the following to the methods:

“To investigate the degree of conservation of CTRP-engaged pathways, we mapped the differentially expressed genes (DEGs) identified from Ctrp10 knockout (KO) versus wild-type (WT) mice to their human orthologs, including human CTRP10, in the GTEx database for transcriptional correlations. Individuals were stratified by sex to examine sex-specific gene connectivity, consisting of 210 males and 100 females to compare gene expression across tissues. Gene-connectivity analyses were performed based on population correlation significances summarized by cumulative -log10(pvalues) as previously described"

Line 330: In Fig 7L, increased oxidative stress in the liver of KO mice is shown. Please provide an explanation for the claim that Ctrp10-KO female mice resembled the WT controls.

In Fig. 7L, we did observe a modest, but significant, increase in oxidative stress in the liver based on the quantification of malondialdehyde (MDA) level, a marker of tissue oxidative stress. However, we did not see any significant differences in the expression of oxidative genes in the liver between WT and KO female mice (Fig. 7J); thus, the statement in line 330 (discussion section) that pertains to oxidative gene expression in fat and liver (Fig. 7E and 7J) is correct.

Line 375: Could you clarify the term "adipose tissue health" and further discuss or provide evidence demonstrating compromised adipose tissue health in female KO mice following HFD?

Adipose tissue health refers to the healthy functioning of adipose tissue (based on its functionality, immune cell population and profile, and metabolic gene expression profiles). Adipose tissue releases free fatty acids in response to fasting and takes up lipids in response to refeeding. Both are these functions are preserved in KO mice as we did not observe any significant differences in free fatty acids (NEFA) and triglyceride levels in the fasted and refed states (Fig. 6AB). Also, we did not observe any significant differences in the expression of inflammatory and fibrotic genes in the adipose tissue of WT and KO female mice fed a high-fat diet (Fig. 7E). If anything, we actually observed a modest, but significant, reduction in the expression of some ER and oxidative stress genes in the KO female mice relative to WT controls (Fig. 7E).

Line 408: Please provide data regarding estrogen levels in wild-type and KO female mice for comparison.

We have now included serum estrogen data. No significant differences in estrogen levels were seen between WT and KO female mice fed either a low-fat diet (Fig. 4 – figure supplement 1) or a high-fat diet (Fig. 5 – figure supplement 2).

Line 587: The GitHub link provided seems to be inactive or incorrect. Please verify and provide the correct link.

We have also updated the github (https://github.com/Leandromvelez/CTRP10-Manuscript-DEG-Sex-specific-connectivities-and-integration) to include a README file and updated the R scripts to annotate steps and processing considerations.

Lines 590-599: Provide additional details about the analysis of human sex-specific genes. Including a table of the top DEGs and pathways differentially regulated by sex would be beneficial for readers' comprehension.

We have expanded the methods, results and associated github repositories to detail all reproducible parameters used in these analyses. The new table of DEGs is included in the manuscript and github repositories.

Version 2

Abstract

Abstract

Obesity is a major risk factor for type 2 diabetes, dyslipidemia, cardiovascular disease, and hypertension. Intriguingly, there is a subset of metabolically healthy obese (MHO) individuals who are seemingly able to maintain a healthy metabolic profile free of metabolic syndrome. The molecular underpinnings of MHO, however, are not well understood. Here, we report that CTRP10/C1QL2-deficient mice represent a unique female model of MHO. CTRP10 modulates weight gain in a striking and sexually dimorphic manner. Female, but not male, mice lacking CTRP10 develop obesity with age on a low-fat diet while maintaining an otherwise healthy metabolic profile. When fed an obesogenic diet, female Ctrp10 knockout (KO) mice show rapid weight gain. Despite pronounced obesity, Ctrp10 KO female mice do not develop steatosis, dyslipidemia, glucose intolerance, insulin resistance, oxidative stress, or low-grade inflammation. Obesity is largely uncoupled from metabolic dysregulation in female KO mice. Multi-tissue transcriptomic analyses highlighted gene expression changes and pathways associated with insulin-sensitive obesity. Transcriptional correlation of the differentially expressed gene (DEG) orthologous in humans also shows sex differences in gene connectivity within and across metabolic tissues, underscoring the conserved sex-dependent function of CTRP10. Collectively, our findings suggest that CTRP10 negatively regulates body weight in females, and that loss of CTRP10 results in benign obesity with largely preserved insulin sensitivity and metabolic health. This female MHO mouse model is valuable for understanding sex-biased mechanisms that uncouple obesity from metabolic dysfunction.

Introduction

The prevalence of obesity has nearly tripled in the past four decades and the underlying cause is complex and multifactorial (1, 2). Genetics, environmental and social factors, and demographics all play a role in contributing to excessive weight gain in the setting of overnutrition (3, 4). Although obesity is a major risk factor for type 2 diabetes, dyslipidemia, cardiovascular disease, and hypertension, not all obese individuals develop the metabolic syndrome (5). There is a subset of metabolically healthy obese (MHO) individuals with an apparently healthy metabolic profile free of some or most components of the metabolic syndrome (6, 7). The molecular and physiological underpinnings of MHO are, however, not well understood. Novel preclinical animal models that can recapitulate features of MHO will be valuable in illuminating pathways that resist the deleterious effects of obesity and provide new therapeutic avenues to mitigate obesity-linked comorbidities.

The mechanisms that normally maintain body weight and metabolic homeostasis are complex and involve both cell autonomous and non-cell autonomous mechanisms. Tissue crosstalk mediated by paracrine and endocrine factors plays an especially important role in coordinating metabolic processes across organ systems to maintain energy balance (8). Of the secretory proteins that circulate in plasma, C1q/TNF-related proteins (CTRP1-15) have emerged as important regulators of insulin sensitivity, and glucose and lipid metabolism (9). We originally identified the first seven members of the CTRP family based on shared sequence homology to the insulin-sensitizing adipokine, adiponectin (10), and subsequently characterized eight additional members (11–16). All fifteen CTRPs share a common C-terminal globular C1q domain and are part of the much larger C1q family (17, 18). The use of gain- and loss-of-function mouse models has helped establish CTRP’s role in controlling various aspects of sugar and fat metabolism (12, 19–34). Additional diverse functions of CTRPs have also been demonstrated in the cardiovascular (35–47), renal (48, 49), immune (20, 50, 51), sensory (52, 53), gastrointestinal (54), musculoskeletal (55–57), and the nervous system (58–61).

Of the family members, CTRP10 (also known as C1QL2) is understudied and consequently only limited information is available concerning its function. The best characterized role of CTRP10 is in the central nervous system (CNS). It has been shown that CTRP10 secreted from mossy fibers is required for the proper clustering of kainite-type glutamate receptors on postsynaptic CA3 pyramidal neurons in the hippocampus (62). It serves as a transsynaptic organizer by directly binding to Neurexin3 (Nrxn3) on the presynaptic terminals, and to glutamate ionotropic receptor kainate type subunit 2 (GluK2) and GluK4 on the postsynaptic terminals (62). Additional putative roles of CTRP10 in the CNS have also been suggested. Genome-wide association studies (GWAS) have implicated CTRP10/C1QL2 in cocaine use disorder (63). In rat models of depression, Ctrp10 expression is increased in the dentate gyrus and reduced in the nucleus accumbens (64). In humans with a history of psychiatric disorders (e.g., schizophrenia), the expression of CTRP10 is elevated in the dorsolateral prefrontal cortex of both males and females (64). Whether and how CTRP10 contributes to addictive behavior and psychiatric disorders is unknown. Recent large-scale efforts to illuminate the genetic architecture of the human plasma proteome also highlighted an association of plasma C1QL2/CTRP10 levels with a missense C4BPA variant (65). In pregnancy, circulating C1QL2/CTRP10 levels are negatively associated with the plasma inflammasome marker, α1-acid glycoprotein (AGP) (66). The biological significance of these clinical observations, however, remains unclear.

The potential function of CTRP10 in peripheral tissues, however, is essentially unknown and unexplored. The present study was motivated by the well documented metabolic functions of many CTRP family members we have characterized to date using genetic loss-of-function mouse models (19–28, 30, 31). We determined that the expression of Ctrp10 in peripheral tissues is modulated by diet and nutritional states, and thus may have a metabolic role. We therefore used a genetic loss-of-function mouse model to determine if CTRP10 is required for regulating systemic metabolism. We unexpectedly discovered a female-specific requirement of CTRP10 for body weight control. We showed that the Ctrp10 KO mice represent a unique female model of MHO with largely preserved insulin sensitivity and metabolic health. This valuable mouse model can be used to inform sex-dependent mechanisms that uncouple obesity from insulin resistance, dyslipidemia, and metabolic dysfunction.

Results

Nutritional regulation of Ctrp10 expression in the brain and peripheral tissues

CTRP10 protein is highly conserved from zebrafish to human (Fig. 1A), with amino acid identity of 67%, 71%, 77%, and 94% between the full-length human protein and the fish, frog, chicken, and mouse orthologs, respectively. The conservation is much higher at the C-terminal globular C1q domain (93-100% identity) between the orthologs. Among the 12 different mouse tissue examined, brain had the highest expression of Ctrp10 (Fig. 1B), consistent with previous findings (67). Expression of Ctrp10 in peripheral tissues was variable and generally much lower than the brain (Fig. 1B). We first determined whether Ctrp10 expression is modulated by nutrition and metabolic state. Male mice (∼10 weeks old) were subjected to fasting and refeeding. In the refed period after an overnight fast, we observed a significant downregulation of Ctrp10 in the visceral (gonadal) white adipose tissue (gWAT), liver, skeletal muscle, kidney, cerebellum, cortex, and hypothalamus relative to the fasted state (Fig. 1C). Next, we examined whether an obesogenic diet alters the expression of Ctrp10. Male mice fed a high-fat diet (HFD) for 12 weeks had a modest increase in Ctrp10 expression in brown adipose tissue (BAT) and heart and decreased expression in skeletal muscle relative to mice fed a control low-fat diet (LFD) (Fig. 1D). These data indicate that Ctrp10 expression is dynamically regulated by acute alterations in energy balance, and perhaps to a lesser extent in alterations to chronic nutritional state on an obesogenic diet.

Nutritional regulation of Ctrp10 expression.

(A) Sequence alignment of full-length human (GenBank # NP_872334), mouse (NP_997116), chicken (XP_046777733), xenopus frog (XP_031749381), and zebrafish (XP_001920705) CTRP10/C1ql2 using Clustal-omega (141). Identical amino acids are shaded black and similar amino acids are shaded grey. Gaps are indicated by dash lines. Signal peptide, collagen domain with characteristic Gly-X-Y repeats, and the C-terminal globular C1q domain are indicated. (B) Ctrp10 expression across different mouse tissues (n = 10). (C) Expression of Ctrp10 across mouse tissues in response to an overnight (16 h) fast or fasting followed by 2 h refeeding. (D) Expression of Ctrp10 across mouse tissues in response to a high-fat diet (HFD) for 12 weeks or a control low-fat diet (LFD). (E) Generation of Ctrp10 knockout (KO) mice. The entire protein coding region in exon 1 and 2 of Ctrp10 was deleted using CRISPR/Cas9 method and confirmed with DNA sequencing. (F) Wild-type (WT) and KO alleles were confirmed by PCR genotyping. (G) The complete loss of Ctrp10 transcript in KO mice was confirmed in mouse cortex, one of the tissues with high Ctrp10 expression (WT, n = 5; KO, n = 5). All expression levels were normalized to β-actin. All data are presented as mean ± S.E.M. * P < 0.05; ** P < 0.01; *** P < 0.001.

Generation of Ctrp10 knockout (KO) mice

We used mice lacking CTRP10 to address whether this secreted protein has a metabolic role in vivo. The mouse Ctrp10 gene consists of two exons (Fig. 1E). The CRISPR-Cas9 method was used to remove the entire protein coding region spanning exon 1 and 2, thus ensuring a complete null allele (Fig. 1E-F). The targeted allele was confirmed by sequencing. As expected, based on the gene deletion strategy, the Ctrp10 transcript was absent from KO mice (Fig. 1G).

CTRP10 is largely dispensable for metabolic homeostasis in young mice fed a control low-fat diet

We first addressed the requirement of CTRP10 in maintaining baseline metabolic homeostasis by assessing metabolic parameters of mice fed a LFD. The body weight and body composition of male mice fed a LFD were not different between genotypes (Fig. 2A-B). By 20 weeks of age, female KO mice fed LFD had a modestly higher body weight relative to WT controls (Fig. 2 C), though the body composition was not different between genotypes (Fig. 2D). Food intake, physical activity, and energy expenditure as measured by indirect calorimetry were also not different between genotypes of either sex across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (Fig. 2E-J). Because Ctrp10 expression is regulated by nutritional states (Fig. 1C), we assessed serum metabolite levels in WT and KO mice in response to fasting and refeeding. No significant differences in fasting and refeeding blood glucose, serum insulin, triglyceride, cholesterol, non-esterified free fatty acids (NEFA), and β-hydroxybutyrate levels were observed between genotypes of either sex, except the female KO mice had slightly lower fasting β-hydroxybutyrate levels (Fig. 3A-B). We performed glucose and insulin tolerance tests to determine any potential differences in glucose handling capacity and insulin sensitivity. No significant differences in glucose and insulin tolerance were noted between genotypes of either sex (Fig. 3C-F). Together, these data indicate that CTRP10 is dispensable for metabolic homeostasis when mice are young (< 20 weeks old) and fed a LFD.

Ctrp10-KO mice fed a low-fat diet have normal body weight and energy balance.

(A-B) Body weight (A) and body composition analysis (B) of fat mass, % fat mass (relative to body weight), lean mass, and % lean mass of WT (n = 17) and KO (n = 14) male mice at 18 weeks of age. (C-D) Body weight (C) and body composition analysis (D) of fat mass, % fat mass (relative to body weight), lean mass, and % lean mass of WT (n = 9) and KO (n = 6) female mice at 13 weeks of age. (E-G) Food intake, physical activity, and energy expenditure (EE) in male mice at 18 weeks of age across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (WT, n = 11-12; KO, n = 10-12). (H-J) Food intake, physical activity, and energy expenditure in female mice at 13 weeks of age (WT, n = 9; KO, n = 6). All data are presented as mean ± S.E.M.

Ctrp10-KO mice fed a low-fat diet have normal fasting-refeeding response and glucose homeostasis.

(A-B) Overnight fasted and refed blood glucose, serum insulin, triglyceride, cholesterol, non-esterified free fatty acids (NEFA), and β-hydroxybutyrate levels in male (A) and female (B) mice. (C-D) Blood glucose levels during glucose tolerance tests (GTT; C) and insulin tolerance tests (ITT; D) in WT (n = 17) and KO (n = 14) male mice at 12 weeks of age. (E-F) Blood glucose levels during glucose tolerance tests (GTT; E) and insulin tolerance tests (ITT; F) in WT (n = 9) and KO (n = 6) female mice at 20 and 21 weeks of age, respectively. All data are presented as mean ± S.E.M. * P < 0.05 (two-way ANOVA with Sidak’s post hoc tests).

CTRP10-deficient female mice on a low-fat diet develop obesity with age

Because female KO mice were slightly heavier at 20 weeks of age (Fig. 2C), we suspected the weight may diverge further with age. Consequently, we monitored the body weight of female mice fed LFD over an extended period. Indeed, the female KO mice gained significantly more weight and adiposity with age (Fig. 4A-C). By the time the mice reached 40 weeks of age, female KO mice weighed ∼ 6 g (20 %) heavier than the WT controls. Consistent with greater adiposity, the adipocyte cell size (cross-sectional area) was also significantly larger in both gonadal (visceral) white adipose tissue (gWAT) and inguinal (subcutaneous) white adipose tissue (iWAT) (Fig. 4D-E). Increased weight gain over time was not attributed to differences in food intake, as measured manually over a 24 h period (Fig. 4F). Fecal output, frequency, and energy content were also not different between genotypes (Fig. 4G), suggesting that weight gain was not due to greater nutrient absorption. Deep colon temperatures in both light and dark cycle were also not different between genotypes (Fig. 4H). Indirectly calorimetry analyses also revealed no significant differences between genotypes in food intake, physical activity, and energy expenditure across the circadian cycle and metabolic states (ad libitum fed, fasted, refed) (Fig. 4I-K). Despite significantly greater body weight and adiposity, female KO mice had the same metabolic profile as the lean WT controls. Given the important role of sex hormone in systemic metabolism, we measured serum estradiol levels and they were not significantly different between WT and KO female mice (Fig. 4 - figure supplement 1). There were no differences in fasting blood glucose, serum insulin, triglyceride, cholesterol, NEFA, and β-hydroxybutyrate levels between genotypes (Fig. 4L). Interestingly, VLDL-TG levels were lower in female KO mice whereas HDL-cholesterol level was not different between genotypes (Fig. 4M). Direct assessments of glucose handling capacity and insulin sensitivity by glucose and insulin tolerance tests, respectively, also revealed no differences between genotypes (Fig. 4N-O). Measurements of mitochondrial respiration in skeletal muscle (gastrocnemius) showed no significant differences in respiration through complex I (CI), CII, and CIV between genotypes; in contrast, Ctrp10 KO female mice had higher respiration at CIV relative to WT controls (Fig. 4 - figure supplement 2). Together, these data indicate that Ctrp10-KO female mice fed LFD develop obesity, but preserve a largely healthy metabolic profile similar to the much leaner WT female mice.

Ctrp10-KO female mice on a low-fat diet develop obesity with age.

(A) Body weights over time of WT and KO female mice fed a low-fat diet (LFD). (B) Representative image of WT and KO female on LFD for 40 weeks. (C) Body composition analysis of WT (n = 9) and KO (n = 6) female mice fed a LFD. (D) Representative H&E stained histology of gonadal white adipose tissue (gWAT) and the quantification of adipocyte cell size (n = 6 per genotype). Scale bar = 100 μM. (E) Representative H&E stained histology of inguinal white adipose tissue (iWAT) and the quantification of adipocyte cell size (n = 6 per genotype). Scale bar = 100 μM. (F) 24-hr food intake data measured manually. (G) Fecal frequency, fecal weight, and fecal energy over a 24 hr period. (H) Deep colon temperature measured at the light and dark cycle. (I-K) Food intake, physical activity, and energy expenditure in female mice across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (WT, n = 9; KO, n = 6). Indirect calorimetry analysis was performed after female mice were on LFD for 30 weeks. (L) Overnight (16-hr) fasted blood glucose, serum insulin, triglyceride, cholesterol, non-esterified free fatty acids, and β-hydroxybutyrate levels. (M) Very-low density lipoprotein-triglyceride (VLDL-TG) and high-density lipoprotein-cholesterol (HDL-cholesterol) analysis by FPLC of pooled (n = 6-7 per genotype) mouse sera. (N) Blood glucose levels during glucose tolerance tests (GTT). (O) Blood glucose levels during insulin tolerance tests (ITT). GTT and ITT were performed when the female mice reached 28 and 29 weeks of age, respectively. WT, n = 9; KO, n = 6.

No significant differences in serum estradiol levels between WT (n = 9) and Ctrp10 KO (n = 6) female mice fed a low-fat diet (LFD).

We included WT male sera (n = 5) as additional controls as chow-fed male mice are expected to have much lower levels of serum estradiol.

Mitochondrial respiration through complex I (CI), CII, and CIV in liver and skeletal muscle (gastrocnemius) of WT (n = 9) and Ctrp10 KO (n = 6) female mice fed a low-fat diet (LFD).

No significant differences in liver and skeletal muscle (gastrocnemius) mitochondrial respiration through complex I (CI), CIII, and CIV between WT (n = 9) and Ctrp10 KO (n = 6) female mice. (A and D) Average oxygen consumption rate (OCR) traces per group, normalized to mitochondrial content. Each group tracing represents the average trace of 9 WT and 6 KO samples. Each tracing shows the entire process of the Seahorse-based respirometry assay with injection compounds listed at the time of introduction to the sample. The sequence is as follows: i) basal reads, ii) addition of NADH (activation of respiration through complex I), iii) addition of antimycin A (AA, inhibitor of complex III) and rotenone (Rot, inhibitor of complex I), iv) addition of TMPD and ascorbate (to activate complex IV via electron donation to cytochrome c), and finally v) addition of azide (inhibitor of complex IV). (B and E) The same information as presented in (A and D) conducted on the same samples, but the NADH injection step is replaced with the injection of succinate (to activate respiration through complex II) and rotenone (to inhibit complex I). (C and F) Average values of all data presented for liver (A-B) and skeletal muscle (D-E). Each dot represents the average of three technical replicates measured at three separate times. Both independent measurements of complex IV (CIV) were used to determine average CIV respiration. * p < 0.05, *** p < 0.001, **** p < 0.0001 (two-way ANOVA with Fisher’s LSD test).

Rapid weight gain in CTRP10-deficient female mice fed a high-fat diet

Next, we challenged the mice with a HFD to determine if the sex-dependent effects on body weight become more pronounced. When fed a HFD, body weight gain and body composition were not different between genotypes in male mice (Fig. 5A-B). Food intake, physical activity, and energy expenditure were also not different between genotypes in male mice across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (Fig. 5C-E). In striking contrast, female KO mice gained weight rapidly on HFD (∼9 g or 28% heavier) and had greater adiposity than the WT controls (Fig. 5F-H). Weight gain in female mice was not due to differences in food intake, body temperature, or changes in fecal output, frequency, and energy content between genotypes (Figure 5 - figure supplement 1). We measured serum estradiol levels and they were not significantly different between WT and KO female mice on HFD (Fig. 5 - figure supplement 2). Metabolic cage analysis also revealed no differences in caloric intake, physical activity, and energy expenditure between genotypes in female mice (Fig. 5I-K). The ANCOVA analysis of energy expenditure using body weight as a covariate also did not reveal any differences between genotypes in female mice (Fig. 5 L). Interestingly, the respiratory quotient (RER) was significantly lower in female KO mice relative to WT controls, especially during fasting and refeeding (Fig. 5M), suggesting a greater reliance on lipid substrates for energy metabolism during those periods. Interestingly, hepatic, but not skeletal muscle, mitochondrial respiration through CII and CIV was lower in Ctrp10 KO female mice relative to WT controls on HFD (Fig. 5 - figure supplement 3). Together, these data indicate that CTRP10 is required for female-specific body weight control in response to caloric surplus, but neither food intake, physical activity level, nor whole-body energy expenditure could account for the marked increase in body weight and adiposity.

Sexually dimorphic response of Ctrp10-KO mice to an obesogenic diet.

(A) Body weights over time of WT and KO male mice fed a high-fat diet (HFD). (B) Body composition analysis of WT (n = 17) and KO (n = 14) male mice fed a HFD for 9 weeks. (C-E) Food intake, physical activity, and energy expenditure in male mice across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed) (WT, n = 11; KO, n = 11). Indirect calorimetry analysis was performed after male mice were on HFD for 10 weeks. (F) Body weights over time of WT and KO female mice fed a high-fat diet. (G) Representative image of WT and KO female mice after 13 weeks of high-fat feeding. (H) Body composition analysis of WT (n = 17) and KO (n = 13) female mice on HFD for 6 weeks. (I-K) Food intake, physical activity, and energy expenditure in female mice (WT, n = 11-12; KO, n = 12) across the circadian cycle (light and dark) and metabolic states (ad libitum fed, fasted, refed). Indirect calorimetry analysis was performed after female mice were on HFD for 6 weeks. (L) ANCOVA analysis of energy expenditure using body weight as a covariate. (M) Respiratory exchange ratio (RER). All data are presented as mean ± S.E.M. * P < 0.05; ** P < 0.01; *** P < 0.001; **** P < 0.0001

No differences in food intake, body temperature, or changes in fecal output, frequency, and energy content between WT and Ctrp10 KO female mice fed a high-fat diet.

Sample size for food intake, fecal frequency, and fecal weight: WT = 6 and KO = 6; sample size for fecal energy and total fecal energy content: WT = 9 and KO = 6; sample size for deep colon temperature: WT = 9 and KO = 10.

No significant differences in serum estradiol levels between WT (n = 6) and Ctrp10 KO (n = 11) female mice fed a high-fat diet (HFD).

We included WT male sera (n = 5) as additional controls as chow-fed male mice are expected to have much lower levels of serum estradiol.

Mitochondrial respiration through complex I (CI), CII, and CIV in liver and skeletal muscle (gastrocnemius) of WT (n = 7) and Ctrp10 KO (n = 8) female mice fed a high-fat diet.

(A and D) Average oxygen consumption rate (OCR) traces per group, normalized to mitochondrial content. Each group tracing represents the average trace of 7 WT and 8 KO samples. Each tracing shows the entire process of the Seahorse-based respirometry assay with injection compounds listed at the time of introduction to the sample. The sequence is as follows: i) basal reads, ii) addition of NADH (activation of respiration through complex I), iii) addition of antimycin A (AA, inhibitor of complex III) and rotenone (Rot, inhibitor of complex I), iv) addition of TMPD and ascorbate (to activate complex IV via electron donation to cytochrome c), and finally v) addition of azide (inhibitor of complex IV). (B and E) The same information as presented in (A and D) conducted on the same samples, but the NADH injection step is replaced with the injection of succinate (to activate respiration through complex II) and rotenone (to inhibit complex I). (C and F) Average values of all data presented for liver (A-B) and skeletal muscle (D-E). Each dot represents the average of three technical replicates measured at three separate times. Both independent measurements of complex IV (CIV) were used to determine average CIV respiration. * p < 0.05, ** p < 0.01 (two-way ANOVA with Fisher’s LSD test).

Obesity is uncoupled from insulin resistance and dyslipidemia in CTRP10-deficient female mice fed a HFD

To evaluate whether the increased adiposity in KO females led to metabolic dysregulation, we again measured fasting and refeeding responses in WT and Ctrp10-KO mice fed a HFD. No differences in fasting and refeeding blood glucose, serum insulin, triglyceride, cholesterol, NEFA, and β-hydroxybutyrate levels were noted between genotypes in male mice (Fig. 6A). Female KO mice, however, had higher fasting blood glucose and serum insulin levels, and lower β-hydroxybutyrate levels compared to the WT controls (Fig. 6B). In the refed state, serum insulin levels continued to be significantly higher in female KO mice. Unlike the WT female mice where refeeding markedly lowered serum β-hydroxybutyrate (ketone) levels as expected, KO female mice appeared unable to suppress serum β-hydroxybutyrate levels in response to refeeding (Fig. 6B).

Ctrp10-KO mice on a high-fat diet have normal glucose and insulin tolerance.

(A-B) Overnight fasted and refed blood glucose, serum insulin, triglyceride, cholesterol, non-esterified free fatty acids (NEFA), and β-hydroxybutyrate levels in male (A) and female (B) mice fed a HFD for 10 weeks. (C-D) Blood glucose levels during glucose tolerance tests (GTT; C) and insulin tolerance tests (ITT; D) in WT (n = 17) and KO (n = 14) male mice fed a HFD for 10 weeks. (E-F) Blood glucose levels during glucose tolerance tests (GTT; E) and insulin tolerance tests (ITT; F) in WT (n = 16) and KO (n = 12) female mice fed a HFD for 8 weeks. (G) VLDL-TG and HDL-cholesterol analysis by FPLC of pooled female mouse sera. All data are presented as mean ± S.E.M. ** P < 0.01; *** P < 0.001; **** P < 0.0001 (two-way ANOVA with Sidak’s post hoc tests for fasted/refed data).

Consistent with the fasting blood glucose and insulin data, glucose handling capacity and insulin sensitivity assessments by glucose and insulin tolerance in HFD-fed male mice also revealed no differences between genotypes (Fig. 6C-D). Female KO mice, however, had higher fasting blood glucose and serum insulin levels suggesting the presence of mild insulin resistance (Fig. 6B). We therefore expected to see differences in either glucose and/or insulin tolerance tests. To our surprise, the rate of glucose clearance in response to glucose or insulin injection was virtually identical between WT and KO female mice (Fig. 6E-F), suggesting no difference in insulin sensitivity between genotypes. VLDL-TG and HDL-cholesterol profiles were also indistinguishable between WT and KO female mice (Fig. 6G). Altogether, these data indicate that CTRP10 is not required for metabolic homeostasis in male mice challenged with a HFD. In female mice, however, loss of CTRP10 markedly promotes weight gain in the face of caloric surplus, but, paradoxically, the excess adiposity is largely uncoupled from obesity-linked insulin resistance and dysregulated glucose and lipid metabolism.

Obesity is uncoupled from adipose dysfunction and hepatic steatosis in Ctrp10-KO female mice fed a HFD

Consistent with greater fat mass in visceral (gonadal) fat depot of Ctrp10-KO female mice (Fig. 7A), histological analysis and quantification also indicated significantly larger adipocyte cell size (Fig. 7B). Although the subcutaneous (inguinal) fat pad weight was also significantly heavier in female KO mice (Fig. 7C), the adipocyte cell size was marginally bigger but not significant (Fig. 7D). A bigger fat pad with only marginally larger cell size suggests greater adipocyte hyperplasia in the subcutaneous depot. Increased adipogenesis in response to caloric surfeit is known to be associated with improved systemic metabolic profile (68).

Ctrp10-KO female mice fed a HFD do not develop adipose tissue dysfunction and fatty liver.

(A) Representative images of dissected gonadal white adipose tissue (gWAT) and the quantification of gWAT weight in WT (n = 15) and KO (n = 12) female mice fed a HFD for 14 weeks. (B) Representative H&E stained histological sections of gWAT and the quantification of adipocyte cell size (n = 7 per genotype). Scale bar = 100 μM. (C) Representative images of dissected inguinal white adipose tissue (iWAT) and the quantification of iWAT weight in WT (n = 15) and KO (n = 12) female mice. (D) Representative H&E stained histological sections of iWAT and the quantification of adipocyte cell size (n = 7 per genotype). Scale bar = 100 μM. (E) Expression of genes associated with inflammation, fibrosis, ER and oxidative stress in gWAT and iWAT of WT (n = 6) and KO (n = 6) female mice fed a HFD for 14 weeks. Gene expression data were obtained from RNA-seq. (F-G) Quantification of hydroxyproline (marker of fibrosis) and malondialdehyde (MDA; marker of oxidative stress) in gWAT and iWAT. WT, n = 15; KO, n = 11. (H) Representative images of dissected liver and the quantification of liver weight in WT (n = 15) and KO (n = 12) female mice. (I) Representative H&E stained histological sections of liver and the quantification of hepatic lipid content (% lipid area; n = 7 per genotype). Scale bar = 100 μM. (J) Hepatic expression of genes associated with inflammation, fibrosis, ER and oxidative stress, lipid synthesis, and lipid catabolism in WT and KO female mice. Gene expression data were obtained from RNA-seq. (K-L) Quantification of hydroxyproline (marker of fibrosis) and malondialdehyde (MDA; marker of oxidative stress) in liver. WT, n = 15; KO, n = 11. All data are presented as mean ± S.E.M. * P < 0.05; ** P < 0.01.

Obesity is known to be associated with low-grade inflammation (69), fibrosis (70), and ER and oxidative stress (71, 72). Despite marked differences in body weight and adiposity, the expression of genes associated with inflammation (except for Ccr2), fibrosis, oxidative stress in gWAT and iWAT were not significantly different between female KO mice and WT controls (Fig. 7E). The expression of some genes associated with ER stress (e.g., Ddit3/CHOP, Atf4, Xbp1) in iWAT were actually lower in female KO mice (Fig. 7E). Corroborating the gene expression data, quantification of hydroxyproline (marker of fibrosis) and malondialdehyde (marker of oxidative stress) revealed no significant differences between genotypes (Fig. 7F-G).

The liver weight of female KO mice was modestly increased (Fig. 7F), but when normalized to body weight it was not significantly different from WT controls (2.76 % in WT and 2.60% in KO, P = 0.24). Histological analysis and quantification revealed no differences in hepatic lipid content (% lipid area) between genotypes (Fig. 7G). Interestingly, although hepatic fat content was similar between genotypes, the expression of lipogenic genes (e.g., Fasn) was lower and fat catabolism genes (e.g., Cpt2, Ppara, Acadl, Acadm, Acad11, Acadvl) was higher in female KO mice (Fig. 7H). The expression of genes associated with inflammation, fibrosis, ER and oxidative stress in liver were not significantly different between genotypes (Fig. 7H). Consistent with the gene expression data, quantification of hydroxyproline (marker of fibrosis) in the liver revealed no significant difference between genotypes (Fig. 7K). The Ctrp10 KO female mice, however, had higher levels of malondialdehyde (a marker of oxidative stress) in the liver, suggesting a modest increase in oxidative stress (Fig. 7L). Altogether, these data indicate that obesity is largely uncoupled from inflammation, fibrosis, ER and oxidative stress in Ctrp10 KO female mice.

Transcriptomic and pathway changes associated with the metabolically healthy obesity phenotype in Ctrp10 KO female mice

To define the specific mechanisms mediating the female-specific effects of Ctrp10 ablation on favorable metabolic outcomes, four major metabolic tissues (gWAT, iWAT, liver, skeletal muscle) from female WT and KO mice fed a HFD were subjected to RNA-sequencing (Figure 8 - Source data 1-8). Comparison of differentially expressed genes (DEGs) via limma (73) showed robust changes across tissues, with the largest changes seen in the liver (Fig. 8A-D). In liver, gWAT, and muscle, we observed comparable numbers of DEGs that were up- and down-regulated, whereas more genes were transcriptionally suppressed in the iWAT of Ctrp10 KO female mice (Fig. 8E, top panel). While significant DEGs were identified in all 4 tissues, only limited overlap was observed between the DEGs in each tissue (Fig. 8E, bottom panel). Gene set enrichment analyses of the DEGs highlighted distinct and shared processes up- or down-regulated across the four tissues (Fig. 8F). Pathways and processes related to lipid metabolism and estrogen receptor were the top-ranked up-regulated enrichments across tissues (Fig. 8F, top panel), whereas processes related to blood clotting and lipoprotein metabolism were the top-ranked down-regulated enrichments (Fig. 8F, bottom panel).

Transcriptomic analysis of liver, adipose tissue, and skeletal muscle of female Ctrp10 KO mice fed a high-fat diet.

(A-D) Cropped volcano plot views of all differentially expressed genes (DEGs, Log2(Fold Change) >1 or <-1 with a p-value <0.05) of the liver, gonadal white adipose tissue (gWAT), inguinal WAT (iWAT), or skeletal muscle (gastrocnemius). (E) Overlap analysis of tissue DEGs showing (top panel) expression unique to gonadal white adipose tissue (gW), inguinal white adipose tissue (iW), liver (L), or skeletal muscle (M). Percent (%) represents percent DEGs unique to each tissue. Bottom panel show DEGs shared across multiple tissues, with all the shared DEGs listed. (F) Enrichr analysis (134) of biological pathways and processes significantly (p<0.01) affected across the CTRP10 deficient female mice. Top pathways and processes derived from Gene Ontology (GO), Reactome (R-HAS), WikiPathway human (WP), and mammalian phenotype (MP). All up- or down-regulated DEGs across all tissues were used for analysis. The tissues contributing to the highest ranked pathways and processes are specified. n = 6 KO and 6 WT for RNA-seq experiments.

Of the DEGs, we found significant changes across tissues in relevant classes of genes that encode proteins involved in gene expression (e.g., transcription factors), signaling (e.g., receptors), tissue crosstalk (e.g., secreted proteins), and metabolism (Fig. 9). Notably, the nuclear receptor, Nr1d1 (also known as Rev-Erbα), is the only gene consistently suppressed across all four tissues (liver, gWAT, iWAT, and muscle) of Ctrp10 KO female mice (Fig. 8E lower panel and Fig. 9). Interestingly, global deletion of Nr1d1 promotes lipogenesis, adipose tissue expansion, and obesity (74, 75). Although the whole-body and adipose-specific Nr1d1 KO mice fed with HFD become markedly obese, the obesity is not accompanied by insulin resistance, adipose tissue inflammation and fibrosis (75, 76). Like the Ctrp10 KO female mice, HFD-fed mice lacking NR1D1 can maintain a relatively healthy metabolic profile despite being strikingly obese. Since only male mice were used in these previous studies, we do not know whether female mice lacking NR1D1 would also exhibit similar insulin-sensitive obesity phenotype. In WT mice, NR1D1 acts as a transcriptional repressor of metabolic genes whose expression are upregulated by high-fat feeding; loss of NR1D1 is thought to result in the de-repression of these genes, leading to greater lipid synthesis and fat mass accrual in response to caloric excess (76). Thus, the suppression of Nr1d1 expression—mimicking NR1D1 deficiency— across tissues in HFD-fed Ctrp10 KO female mice may contribute to benign fat mass expansion without the accompanying adipose tissue fibrosis, inflammation, and oxidative stress.

Loss of CTRP10 induces significant and wide-spread alterations in the expression of key transcription factors, secreted protein, membrane receptors, and metabolism-associated genes.

(A-D) Selected genes from the DEG list of each tissue organized based on gene type (genes encoding transcription factors, secreted proteins, receptors, and proteins involved in metabolism) and ranked from highest to lowest average row z-score. Z-score is defined as z = (x-μ)/σ, where x is the raw score (gene transcript level), μ is the population mean (i.e., mean of gene expression across both WT and KO samples), and σ is the population standard deviation. n = 6 per genotype.

While hepatic Fgf21 expression is significantly elevated, its systemic circulating levels were only found to be higher (P = 0.051) in LFD-fed, but no HFD-fed, KO female mice.

Among the upregulated genes, Fgf21 and Fgf1 were significantly elevated in the liver of Ctrp10 KO female mice (Fig. 9A). FGF21 is an hepatokine known to improve systemic insulin sensitivity and to promote a favorable metabolic profile in diet-induced obese mice (77). While hepatic Fgf21 expression is significantly elevated, its systemic circulating levels were only found to be higher (P = 0.051) in LFD-fed, but not HFD-fed, KO female mice (Fig. 9 – figure supplement 1); this suggests liver-derived FGF21 may be retained and act locally within the liver parenchyma. FGF1 is another secreted protein that has been shown to dampen hepatic glucose output by suppressing adipose lipolysis (78), improve systemic insulin sensitivity by reducing adipose inflammation (79), and alleviate hepatic steatosis, inflammation, and insulin resistance (80). Thus, upregulated expression of Fgf21 and Fgf1 in Ctrp10 KO female mice could contribute to the MHO phenotype. In addition, the upregulated hepatic expression of IL-22 receptor (Il22ra1) in Ctrp10 KO female mice (Fig. 9A) may confer protection against obesity-associated fatty liver, inflammation and fibrosis (81–83). Further, a marked increase in uncoupling protein 3 (Ucp3) and Krüppel-like factor 15 (Klf15) expression in the skeletal muscle (Fig. 9C) may promote lipid utilization and help mitigate lipid-induced insulin resistance in Ctrp10 KO female mice (84, 85). Taken together, these combined changes—at the level of gene expression and biological pathways and processes across tissues—acting in concert likely contribute to the apparently healthy obesity phenotype seen in the KO female mice.

Conservation of mouse DEG co-correlation in humans highlights sex-specific gene connectivity

Next, we asked whether the female-specific transcriptomic effects across tissues were conserved in humans. To address this, we analyzed transcriptional co-correlation of mouse DEG (Fig. 10) orthologues in GTEx (86), consisting of 210 males and 100 females filtered for comparison of gene expression across tissues (87, 88). Hierarchical clustering of transcriptional correlation of the orthologous DEGs among 4 metabolic tissues— subcutaneous and visceral white adipose tissue, liver, and skeletal muscle—showed differing patterns of gene connectivity between females (Fig. 10A) and males (Fig. 10B). When grouped according to sex in each tissue, the degree of sex-specific gene correlation pairs of DEGs orthologues showed the most significant differences in subcutaneous adipose tissue (Fig. 10C). Given the whole-body metabolic effects of Ctrp10 ablation in mice, we further examined the degree of sex-dependent DEG co-correlation across metabolic tissues. This analysis showed that human orthologue genes in subcutaneous adipose tissue (Fig. 10D, top row) and liver (Fig. 10D, third row) also exhibited highly significant sex differences in their transcriptional correlation with other DEGs across key metabolic tissues (Fig. 10D). These analyses highlight the sex-specificity of CTRP10 DEG orthologues in humans, suggest possible sex-biased mechanisms of tissue crosstalk, and overall underscores the conservation of the sex-dependent metabolic function of CTRP10.

GTEx genetic co-correlation of mouse differentially expressed gene (DEG) orthologues.

(A-B) Heatmaps showing biweight midcorrelation (bicor) coefficient among human tissue DEG orthologues in females (A) and males (B) in GTEx. Y-axis color indicates tissue of origin, P-value based on students’ regression P-value. (C) T-tests between correlation coefficient in males and females among all DEG orthologue gene pairs for subcutaneous (SubQ) adipose tissue, visceral (visc) adipose tissue, liver, and skeletal muscle. (D) the same as in C, except comparisons are shown for all gene-gene pairs between tissues. For example, the top left graph compared the connectivity of males (blue color) vs females (green color) for correlation between subcutaneous (SubQ) and visceral (Visc) adipose tissue DEG orthologues.

Discussion

Our current study has established a novel function for CTRP10 in modulating body weight in a sex-specific manner. When mice were fed a control LFD, female Ctrp10-KO mice developed obesity with age; increased adiposity, however, did not impair insulin action and glucose and lipid metabolism. When challenged with an obesogenic diet, female Ctrp10-KO mice gained weight rapidly. Despite having strikingly higher adiposity and weighing ∼10-11 g (∼28%) more, female KO mice fed a HFD exhibited a metabolic profile largely indistinguishable from the much leaner WT controls. Although female KO mice had higher fasting glucose and insulin levels, direct assessments of glucose metabolism and insulin sensitivity by glucose and insulin tolerance tests, however, revealed no differences between genotypes. Except having lower fasting ketone (β-hydroxybutyrate) levels, the fasting lipid profile, as well as VLDL-TG and HDL-cholesterol levels, of Ctrp10-KO female mice resembled the WT controls. The hepatic fat content was also comparable between genotypes. Global transcriptomic profiling across different fat depots and liver did not reveal gene expression signatures associated with elevated inflammation, fibrosis, and ER and oxidative stress. Altogether, these findings suggest that CTRP10 deficiency promotes obesity in females but it also uncouples obesity from insulin resistance, dyslipidemia, steatosis, inflammation, and oxidative stress. Thus, Ctrp10-KO female mice represent a novel model of female obesity with largely preserved insulin sensitivity and metabolic health.

Our findings help inform ongoing studies on metabolically healthy obese (MHO) humans (89–94). Because the criteria used to define MHO differs between studies, there is an ongoing debate regarding the prevalence of MHO and what fraction of the MHO population is insulin-sensitive and metabolically healthy (95, 96). Nevertheless, among the obese individuals, there clearly exists a subgroup that maintains long-term normal insulin sensitivity and does not appear to develop any component of the metabolic syndrome (95). MHO is observed in both sexes, but more common in females (6). The underlying mechanism(s) that uncouple obesity from adverse metabolic health in MHO is not well understood (7, 91, 95). It is currently unknown whether males and females with MHO use similar or distinct mechanism to maintain insulin sensitivity and metabolic health. Our findings in Ctrp10-KO female, but not male, mice suggest that there may be female-biased mechanism that prevents metabolic deterioration in the face of obesity, thus underscoring the utility of the Ctrp10-KO mice as a female mouse model of MHO.

Obesity is frequently associated with insulin resistance, dyslipidemia, fatty liver, oxidative stress, and chronic low-grade inflammation (69, 97, 98). The mechanisms that link obesity to metabolic dysfunctions are complex and multifactorial. There are limited number of mouse models described where obesity is uncoupled from insulin resistance and metabolic health (99–102); in some studies, however, only male mice were used or that the sex of the animals was not specified. In the case of aP2/FABP4 KO male mice, the uncoupling of obesity from insulin resistance was attributed to a marked decrease in TNF-α expression in adipose tissue (100). In the case of adiponectin overexpression in leptin-deficient (ob/ob) male and female mice, a dramatic expansion of the subcutaneous fat pad is thought to promote lipid sequestration in adipose compartment, thus preventing ectopic lipid deposition in non-adipose tissues (e.g., liver, pancreas, muscle) that would otherwise induce insulin resistance (99). Massive obesity with preserved insulin sensitivity is also observed in leptin-deficient (ob/ob) male and female mice overexpressing the mitochondrial membrane protein, mitoNEET (102). The benign obesity is attributed to the inhibition of iron transport into mitochondria by mitoNEET, leading to reduced mitochondrial activity, fatty acid oxidation, and oxidative stress (102). In the Brd2 hypomorphic mice, severe obesity with lower blood glucose and enhanced glucose tolerance is due to a combination of hyperinsulinemia and marked reduction in macrophage infiltration into fat depot (103). Lastly, in male mice fed a high starch diet, the uncoupling of obesity from insulin resistance is associated with lower ceramide levels in liver and skeletal muscle (101). In all these cases, the uncoupling of obesity from metabolic dysfunction is seen in either male mice only (female mice were not included) or both sexes. These previous studies suggest that multiple mechanisms, not mutually exclusive, can contribute to the MHO phenotype in different mouse models.

In our study, loss of CTRP10 in female mice largely uncoupled obesity from insulin resistance, dyslipidemia, steatosis, inflammation, and ER and oxidative stress. The preservation of insulin sensitivity in Ctrp10 KO female mice is due, at least in part, to the absence of obesity-linked adipose and liver inflammation, fibrosis, and oxidative stress. These phenotypes typically associated with a favorable metabolic profile are also observed in MHO individuals (90, 104). A healthy adaptive remodeling of white adipose tissues in response to caloric surfeit helps preserve the storage and secretory function of adipocytes (105). The expansion of benign adipose tissues further serves to sequester circulating lipids and prevent their ectopic deposition in non-adipose tissue (e.g., liver and skeletal muscle) which can impair insulin action (106, 107). The MHO phenotype seen in Ctrp10 KO female mice reinforce the notion that adipose tissue health, rather than abundance, is an important determinant of metabolic health in obesity.

Because lipidomic analysis was not performed—a limitation of this study—we do not know whether Ctrp10- KO female mice have reduced ceramide or diacylglycerol levels in liver and skeletal muscle, two lipid species known to antagonize insulin action (108, 109). However, our global transcriptomic and pathway enrichment analysis across visceral and subcutaneous fat depots, liver, and skeletal muscle highlighted the relevant up- and down-regulated pathways and processes (e.g., lipid and lipoprotein metabolism, signaling) that may contribute to the MHO phenotype in Ctrp10-KO female mice. How these changes across tissues help to suppress the deleterious effects of obesity and maintain an apparently healthy metabolic profile in Ctrp10-KO female mice remains to be fully elucidated. Part of the mechanism may be attributable to the suppression of Nr1d1 and the upregulated expression of Fgf1, Fgf21, Il22ra1, Ucp3, Klf15. Increased expression of these genes is known to reduce obesity-linked inflammation, oxidative stress, steatosis, and insulin resistance. Interestingly, global deletion of Nr1d1 preserves insulin sensitivity despite promoting lipogenesis, adipose tissue expansion, and obesity (74, 75). NR1D1 (also known as REV-ERBα) is a nuclear receptor that acts as a transcriptional repressor (110, 111). Circadian and metabolic genes that normally repressed by NR1D1 are de-repressed in mice lacking NR1D1 globally or in a fat-specific manner, thus leading to fat mass accrual (74, 76). It remains to be determined whether the suppression of Nr1d1 expression in adipose tissue, liver, and skeletal muscle is indeed responsible, at least in part, for the MHO phenotype seen in Ctrp10 female mice.

In many single-gene KO mouse models where both sexes are examined, it is often the males that show a more pronounced metabolic phenotype. It is known that C57BL/6J female mice generally gain significantly less weight on HFD compared to male mice (112). Therefore, it is intriguing that Ctrp10-KO female mice became obese on a control LFD and gained weight rapidly when fed an obesogenic diet. After twelve weeks on HFD, the body weight of female KO mice was approaching that of WT male mice fed the same diet. What mechanism underlies the sexually dimorphic requirement of CTRP10 for body weight control? We know that the obesity phenotype was not attributed to differences in food intake, physical activity level, body temperature, and energy expenditure between WT and KO female mice. We assume that the methods used to quantify these physiologic parameters are sensitive enough to detect small differences that can give rise to divergent body weight over time. Quantification of fecal output and fecal energy content also revealed no differences between genotypes. Thus, loss of CTRP10 did not affect macronutrient intake and absorption. Although we cannot fully rule out the CNS function of CTRP10/C1QL2 (62), our data do not support a central role for CTRP10 in modulating food intake behavior, locomotor activity, and energy expenditure that affect body weight in female mice.

It is known that reduced estrogen level by ovariectomy or blocking estrogen action in estrogen receptor (ERα) KO mice will cause obesity and metabolic dysfunction in female mice fed a HFD (113–115). Conversely, estradiol supplementation decreases HFD-induced weight gain and improves glucose tolerance and insulin sensitivity (116–118). Estrogen also has the effect of reducing food intake, and promoting physical activity and energy expenditure (119–123). In our study, loss of CTRP10 promotes obesity without altering food intake, physical activity, and energy expenditure. While estrogen’s role cannot be completely ruled out, the fact that estrogen levels were not different between genotypes and that female Ctrp10-KO mice developed obesity with largely preserved metabolic health suggest that factors other than altered estrogen level contribute to the insulin-sensitive obesity phenotype. Future studies are warranted to uncover what factor(s) is causally contributing to obesity in female mice lacking CTRP10.

The sex-dependent effects of CTRP10 on metabolism and tissue transcriptomes appear to be conserved in humans. When the human orthologues of the mouse DEGs were used to interrogate the GTEx data, clear patterns of gene connectivity within and across metabolic tissues in females and males were observed, with the strongest sex-specific gene correlations seen in subcutaneous adipose tissue and liver. These findings provide further evidence that CTRP10 modulates tissue transcriptome in a sex-dependent manner. Further, our analyses of sex-dependent DEG co-correlation across metabolic tissues also suggest possible sex-biased mechanisms of inter-organ metabolic signaling between adipose tissue and liver.

CTRP10 has been previously shown to bind to the adhesion GPCR, brain angiogenesis inhibitor-3 (Bai3/Adgrb3) (124). Bai3 is expressed in the brain and peripheral tissues, and it is a promiscuous GPCR that can bind to multiple ligands. In addition to CTRP10 (C1QL2), Bai3 also binds to CTRP11 (C1QL4), CTRP13 (C1QL3), CTRP14 (C1QL1), neuronal pentraxins, and reticulon 4 (RTN4) receptor (56, 58, 59, 124–126). A constitutive, whole-body KO of Bai3 mouse models have recently been generated (127, 128). Both male and female Bai3 KO mice fed a standard chow have significantly lower body weight, beginning at weaning (3 weeks old) and continue into adulthood (127, 128). Lower body weight in Bai3 KO mice of either sex is attributed to a reduction in both lean and fat mass, and is associated with higher energy expenditure and reduced food intake in male mice (127). The impact of Bai3 deficiency on systemic metabolism in response to a high-fat diet was not examined. The Ctrp10 KO mice do not phenocopy the phenotypes of the Bai3 KO mice. When fed a control low-fat diet, the body weight, food intake, and energy expenditure of Ctrp10 KO male mice were indistinguishable from WT controls. In striking contrast to Bai3 KO mice, female Ctrp10 KO mice fed a low-fat diet began to gain more weight around 20 weeks of age, and by 40 weeks had become visibly obese. While we do not rule out CTRP10-Bai3 signaling axis in modulating energy metabolism in peripheral tissues, our findings in Ctrp10 KO mice suggest that future works are needed to establish the molecular mechanisms that mediate the systemic metabolic function of CTRP10.

Several limitations of our current study are noted. The lack of reliable antibody specific for CTRP10 preclude the assessment of protein abundance in response to altered nutritional and metabolic states. We used a constitutive whole-body KO mouse model of CTRP10 to interrogate its function. It is unknown whether CTRP10 has a role during development that may influence sex-dependent postnatal weight gain with age or in response to a high-caloric diet. Future studies using conditional KO of Ctrp10 gene in adult mice can help address this issue. We also did not perform tracer studies to address whether there are anabolic and catabolic changes in glucose and fat metabolism in the liver and adipose tissue of Ctrp10 KO female mice, as this information may shed light on how different tissues handle macronutrients in the absence of CTRP10. In KO female mice, we observed upregulated expression of genes (e.g., Fgf1, Fgf21, Il22ra1, Ucp3, and Klf15) in liver or skeletal muscle that potentially contributes to the MHO phenotype. The mechanism by which CTRP10 suppresses—either directly or indirectly—the expression of these genes and their de-repression in the absence of CTRP10 is presently unknown. Although the lack of differences in food intake, physical activity, and energy expenditure between genotypes do not support a central role of CTRP10 in mediating the metabolic phenotypes of Ctrp10-KO female mice, a brain-specific KO of Ctrp10 gene is needed to definitively rule this out. Our phenotypic analyses in the context of metabolism are relatively comprehensive but not exhaustive. Although the metabolic profile of obese Ctrp10-KO female mice was largely indistinguishable from the much leaner WT controls, some related aspect of metabolic health (e.g., blood pressure and heart function) may be altered in the absence of CTRP10 which we did not examine.

In summary, we have established the physiologic role and requirement of CTRP10 in modulating body weight in a female-specific manner. Importantly, loss of CTRP10 uncouples obesity from insulin resistance and metabolic dysfunction. The CTRP10-deficient female mice represent a unique and valuable model for dissecting female-biased mechanisms that help preserve metabolic health in the face of positive energy balance and increased adiposity.

Materials and methods

Mouse models

Eight-week-old mouse tissues (gonadal and inguinal white adipose tissues, interscapular brown adipose tissue, liver, heart, skeletal muscle, kidney, pancreas, cerebellum, cortex, hippocampus, hindbrain, and hypothalamus) from C57BL/6J male mice (The Jackson Laboratory, Bar Harbor, ME) were collected from fasted and refed experiments as we have previously described (129). For the fasted group (∼10 weeks old), food was removed for 16 h (beginning 10 h into the light cycle), and mice were euthanized 2 h into the light cycle. For the refed group (∼10 weeks old), mice were fasted for 16 h and refed with chow pellets for 2 h before being euthanized. Tissues (white and brown adipose tissues, liver, whole brain, kidney, spleen, heart, skeletal muscle, pancreas, small intestine, and colon) from C57BL/6J male mice fed a low-fat diet (LFD) or a high-fat diet (HFD) for 12 weeks were also collected as we have previously described (129).

The Ctrp10/C1ql2-KO mice (C57BL/6NCrl-C1ql2em1(IMPC)Mbp/Mmucd; stock number 050587-UCD) were generated using the CRISPR-cas9 method at UC Davis. The two guide RNAs (gRNA) used were 5’-CCGGCGCC GCTCCACCATTACCT-3’ and 5’-TCAGGCCACCCCATCCCCATCGG-3’. The Ctrp10 gene consists of 2 exons. The entire protein coding region spanning exon 1 and 2 was deleted. This KO strategy ensures a complete null allele for Ctrp10. The KO mice were maintained on a C57BL6/6J genetic background. Genotyping primers for WT allele were forward (m10-Com-F) 5’-TGTCGGGCTCTTCGACTCTCCA-3’ and reverse (m10-WT-R) 5’-GCATCTCGT AGTGAGCCGCTCC-3’. The size of the WT band was 360 bp. Genotyping primers for the Ctrp10 KO allele were forward (m10-Com-F) 5’-TGTCGGGCTCTTCGACTCTCCA-3’ and reverse (m10-Mut-R1) 5’-GTCCAATCAGCT TTCTCAAGTCTGG-3’. The size of the KO band was 422 bp. The genotyping PCR parameters were as follows: 94°C for 5 min, followed by 10 cycles of (94°C for 10 sec, 65°C for 15 sec, 72°C for 30 sec), then 25 cycles of (94°C for 10 sec, 55°C for 15 sec, 72°C for 30 sec), and lastly 72°C for 5 min. Due to the presence of GC rich sequences, 7% DMSO was included in the PCR genotyping reaction. Mice were generated by intercrossing Ctrp10 heterozygous (+/-) mice, supplemented with intercrossing WT or KO mice. Ctrp10 KO (-/-) and WT (+/+) controls were housed in polycarbonate cages on a 12-h light–dark photocycle with ad libitum access to water and food. Mice were fed either a control low-fat diet (LFD; 10% kcal derived from fat; # D12450B; Research Diets, New Brunswick, NJ) or a high-fat diet (HFD; 60% kcal derived from fat; #D12492, Research Diets). LFD was provided for the duration of the study, beginning at 5 weeks of age; HFD was provided for 14 weeks, beginning at 6-7 weeks of age. At termination of study, all mice were fasted for 2 h and euthanized. Tissues were collected, snap-frozen in liquid nitrogen, and kept at −80°C until analysis. All mouse protocols (protocol # MO22M367) were approved by the Institutional Animal Care and Use Committee of the Johns Hopkins University School of Medicine. All animal experiments were conducted in accordance with the National Institute of Health guidelines and followed the standards established by the Animal Welfare Acts.

Body composition analysis

Body composition analyses for total fat, lean mass, and water content were determined using a quantitative magnetic resonance instrument (Echo-MRI-100, Echo Medical Systems, Waco, TX) at the Mouse Phenotyping Core facility at Johns Hopkins University School of Medicine.

Indirect calorimetry

LFD- or HFD-fed WT and Ctrp10 KO male and female mice were used for simultaneous assessments of daily body weight change, food intake (corrected for spillage), physical activity, and whole-body metabolic profile in an open flow indirect calorimeter (Comprehensive Laboratory Animal Monitoring System, CLAMS; Columbus Instruments, Columbus, OH) as previously described (25). In brief, data were collected for three days to confirm mice were acclimatized to the calorimetry chambers (indicated by stable body weights, food intakes, and diurnal metabolic patterns), and data were analyzed from the fourth day. Rates of oxygen consumption (O2; mL·kg-1·h-1) and carbon dioxide production (CO2; mL·kg-1·h-1) in each chamber were measured every 24 min throughout the studies. Respiratory exchange ratio (RER CO2/O2) was calculated by CLAMS software (version 5.66) to estimate relative oxidation of carbohydrates (RER = 1.0) versus fats (RER = 0.7), not accounting for protein oxidation. Energy expenditure (EE) was calculated as EE= O2× [3.815 + (1.232 × RER)] and normalized to lean mass. Because normalizing to lean mass can potentially lead to overestimation of EE, we also performed ANCOVA analysis on EE using body weight as a covariate (130). Physical activities were measured by infrared beam breaks in the metabolic chamber.

Measurements of 24 h food intake

To independently confirm the food intake data collected in the metabolic cage (CLAMS), we also performed 24 h food intake measurements manually. All mice were singly housed, with wire mesh flooring inserts over a piece of cage paper on the bottom of the cage. A known weight of food pellets was given to each mouse. Twenty-four hours later, the leftover food pellets remaining on the flooring insert, along with any spilled crumbs on the cage paper, were collected and weighed. Thus, food intake was corrected for spillage.

Glucose, insulin, pyruvate, and lipid tolerance tests

All tolerance tests were conducted as previously described (21, 24, 29). For glucose tolerance tests (GTTs), mice were fasted for 6 h before glucose injection. Glucose (Sigma, St. Louis, MO) was reconstituted in saline (0.9 g NaCl/L), sterile-filtered, and injected intraperitoneally (i.p.) at 1 mg/g body weight (i.e., 10 μL/g body weight). Blood glucose was measured at 0, 15, 30, 60, and 120 min after glucose injection using a glucometer (NovaMax Plus, Billerica, MA). For insulin tolerance tests (ITTs), food was removed 2 h before insulin injection. 6.5 μL of insulin stock (4 mg/mL; Gibco) was diluted in 10 mL of saline, sterile-filtered, and injected i.p. at 0.75 U/kg body weight (i.e., 10 μL/g body weight). Blood glucose was measured at 0, 15, 30, 60, and 90 min after insulin injection using a glucometer (NovaMax Plus).

Fasting-Refeeding insulin tests

Mice were fasted overnight (∼16 h) then reintroduced to food as described (26). Blood glucose was monitored at the 16 h fast time point (time = 0 h refed) and at 1 and 2 hours into the refeeding process. Serum was collected at the 16 h fast and 2 h refed time points for insulin ELISA, as well as for the quantification of triglyceride, cholesterol, non-esterified free fatty acids (NEFA), and β-hydroxybutyrate levels.

Blood and tissue chemistry analysis

Tail vein blood samples were allowed to clot on ice and then centrifuged for 10 min at 10,000 x g. Serum samples were stored at −80°C until analyzed. Serum triglycerides (TG) and cholesterol levels were measured according to manufacturer’s instructions using an Infinity kit (Thermo Fisher Scientific, Middletown, VA). Non-esterified free fatty acids (NEFA) were measured using a Wako kit (Wako Chemicals, Richmond, VA). Serum β-hydroxybutyrate (ketone) concentrations were measured with a StanBio Liquicolor kit (StanBio Laboratory, Boerne, TX). ELISA kits were used to measure serum insulin (Crystal Chem, Elk Grove Village, IL; cat # 90080), estradiol (Calbiotech, El Cajon, CA; cat # ES380S), and FGF21 (R&D Systems, Minneapolis, MN; cat # MF2100) according to the manufacturer’s instructions. Hydroxyproline assay (Sigma Aldrich, MAK008) was used to quantify total collagen content in liver and adipose tissues according to the manufacturer’s instructions. Lipid peroxidation levels (marker of oxidative stress) in the liver and adipose tissues were assessed by the quantification of malondialdehyde (MDA) via Thiobarbituric Acid Reactive Substances (TBARS) assay (Cayman Chemical, 700870) according to the manufacturer’s instructions.

Serum lipoprotein triglyceride and cholesterol analysis by FPLC

Food was removed for 2-4 hr (in the light cycle) prior to blood collection. Sera collected from mice were pooled (n = 6-7/genotype) and sent to the Mouse Metabolism Core at Baylor College of Medicine for analysis. Serum samples were first fractionated by fast protein liquid chromatography (FPLC). A total of 45 fractions were collected, and TG and cholesterol in each fraction was quantified.

Histology and quantification

Inguinal (subcutaneous) white adipose tissue (iWAT), gonadal (visceral) white adipose tissue (gWAT), and liver were dissected and fixed in formalin. Paraffin embedding, tissue sectioning, and staining with hematoxylin and eosin were performed at the Pathology Core facility at Johns Hopkins University School of Medicine. Images were captured with a Keyence BZ-X700 All-in-One fluorescence microscope (Keyence Corp., Itasca, IL). Adipocyte (gWAT and iWAT) cross-sectional area (CSA), as well as the total area covered by lipid droplets in hepatocytes were measured on hematoxylin and eosin-stained slides using ImageJ software (131). For CSA measurements, all cells in one field of view at 100X magnification per tissue section per mouse were analyzed. Image capturing and quantifications were carried out blinded to genotype.

Fecal bomb calorimetry and assessment of fecal parameters

Fecal pellet frequency and average fecal pellet weight were monitored by housing each mouse singly in clean cages with a wire mesh sitting on top of a cutout cardboard that lay at the bottom of the cage for fecal collection. The number of fecal pellets and their total weight was recorded at the end of 24 h period. Additional fecal pellets collected for 3 full days were combined and shipped to the University of Michigan Animal Phenotyping Core for fecal bomb calorimetry. Briefly, fecal samples were dried overnight at 50°C prior to weighing and grinding them to powder. Each sample was mixed with wheat flour (90% wheat flour, 10% sample) and formed into 1.0 g pellet, which was then secured into the firing platform and surrounded by 100% oxygen. The bomb was lowered into a water reservoir and ignited to release heat into the surrounding water. These data were used to calculate fecal pellet frequency (bowel movements/day), average fecal pellet weight (g/bowel movement), fecal energy (cal/g feces), and total fecal energy (kcal/day).

Mitochondrial respirometry of tissue samples

High-resolution respirometry was conducted on previously frozen tissue samples (liver and gastrocnemius muscle) to assay for mitochondrial activity as we have previously described (132). Briefly, all tissues were dissected, snap frozen in liquid nitrogen, and stored at −80°C for later analysis. Samples were thawed in MAS buffer (70mM sucrose, 220 mM mannitol, 5 mM KH2PO4, 5 mM MgCl2, 1 mM EGTA, 2 mM HEPES pH 7.4), finely minced with scissors, and then homogenized with a glass Dounce homogenizer. The resulting homogenate was spun at 1000 g for 10 min at 4°C. The supernatant was collected and immediately used for protein quantification by BCA assay (Thermo Scientific, 23225). Each well of the Seahorse microplate was loaded with 8 µg (liver samples) or 10 µg (gastrocnemius muscle samples) of homogenate protein. Each biological replicate is comprised of three technical replicates. Samples from all tissues were treated separately with NADH (1 mM) as a complex I substrate or Succinate (a complex II substrate, 5 mM) in the presence of rotenone (a complex I inhibitor, 2 µM), then with the inhibitors rotenone (2 µM) and Antimycin A (4 µM), followed by TMPD (0.45 mM) and Ascorbate (1 mM) to activate complex IV, and finally treated with Azide (40 mM) to assess non-mitochondrial respiration. All respiration data were normalized to mitochondrial content, quantified using MitoTracker Deep Red (MTDR, ThermoFisher, M22426) as described (132). Briefly, lysates were incubated with MTDR (1 µM) for 10 min at 37°C, then centrifuged at 2000 g for 5 min at 4°C. The supernatant was carefully removed and replaced with 1x MAS solution and fluorescence was read with excitation and emission wavelengths of 625 nm and 670 nm, respectively. To minimize non-specific background signal contribution, control wells were loaded with MTDR and 1x MAS and subtracted from all sample values.

Tissue library preparation and RNA sequencing

Total RNA was isolated from tissues using Trizol reagent (Thermo Fisher Scientific) according to the manufacturer’s instructions. Library preparation and bulk RNA sequencing of liver, skeletal muscle (gastrocnemius), gonadal white adipose tissue (gWAT), and inguinal white adipose tissue (iWAT) of HFD-fed Ctrp10-KO female mice and WT controls were performed by Novogene (Sacramento, California, USA) on an Illumina platform (NovaSeq 6000) and pair-end reads were generated. Sample size: 6 WT and 6 KO for each tissue. All raw sequencing files are available from the NIH Sequence Read Archive (SRA) accession PRJNA971939.

Mouse RNA-Sequencing analysis

Transcript features were assembled from raw fastq files and aligned to the current version of mouse transcriptome (Mus musculus.GRCm39.cdna) using kallisto-aln (133). Version-specific Ensembl transcript IDs were linked to gene symbols using biomart. Estimated counts were log normalized and filtered for a limit of sum >5 across all samples. Logistic regressions comparing WT vs KO samples across tissues were performed using limma (73). Differential expression results were visualized using available R packages in CRAN: ggplot2, ggVennDiagram and pheatmap. Gene set enrichment analyses of the DEGs were performed using Enrichr (134). Scripts for analyses and visualization are available at the github link below (Data availability section).

Human sex difference analysis

All the datasets and scripts to perform analyses are available at the github link below (Data availability section). Male and Female human data were obtained from Genotype-Tissue Expression (GTEx) (86) and filtered for sufficient comparison of inter-tissue transcript correlation as described (87, 88). CTRP10/C1QL2 and other human orthologues for mouse differentially expressed genes (DEGs) were identified by intersecting mouse gene symbols with known human orthologues from the vertebrate homology resource at Mouse Genome Informatics (MGI) (135). Co-correlation between all human orthologues DEGs were calculated in either self-reported male or female subjects in GTEx using the bicorAndPvalue() function in Weighted Genetic Coexpression Network Analysis (WGCNA) package (136). To compare sex-differences of regression coefficients, Wilcoxon t-tests were compared between coefficients using the R package ggpubr. To investigate the degree of conservation of CTRP-engaged pathways, we mapped the DEGs identified from Ctrp10 KO versus WT mice to their human orthologs, including human CTRP10/C1QL2, in the GTEx database for transcriptional correlations. Individuals were stratified by sex to examine sex-specific gene connectivity, consisting of 210 males and 100 females to compare gene expression across tissues. Gene-connectivity analyses were performed based on population correlation significances summarized by cumulative -log10(pvalues) as previously described (87, 137–139). Additional details on the filtering criteria and normalization methods can be found in the GitHub repository.

Quantitative real-time PCR

Total RNA was isolated from tissues using Trizol reagent (Thermo Fisher Scientific). Purified RNA was reverse transcribed using an iScript cDNA Synthesis Kit (Bio-rad). Real-time quantitative PCR analysis was performed on a CFX Connect Real-Time System (Bio-rad) using iTaqTM Universal SYBR Green Supermix (Bio-rad) according to manufacturer’s instructions. Data were normalized to the stable housekeeping gene β-actin or 36B4 (encoding the acidic ribosomal phosphoprotein P0) and expressed as relative mRNA levels using the ΔΔCt method (140). Real-time qPCR primers used to assess Ctrp10 expression across mouse tissues were: Ctrp10 forward, 5’-CGGCTTCATGAC ACTTCCTGA-3’ and reverse, 5’-AGCAGGGATGTGTCTTTTCCA-3’. qPCR primers used to confirm the absence of Ctrp10 in KO mice were: forward (qPCR-m10-F2), 5’-CACGTACCACATTCTCATGCG-3’ and reverse (qPCR-m10-R1), 5’-TCGTAATTCTGGTCCGCGTC-3’.

Statistical analyses

Sample size is indicated in figure and/or figure legend. All results are expressed as mean ± standard error of the mean (SEM). Statistical analysis was performed with Prism 9 software (GraphPad Software, San Diego, CA). Data were analyzed with two-tailed Student’s t-tests, one-way ANOVA or two-way ANOVA (with Sidak’s post hoc tests). 2-way ANOVA was used for body weight over time, fasting-refeeding response, and all tolerance tests. P < 0.05 was considered statistically significant.